### 2016-05-25 paper.Rmd
yo-yo

---
title: "Automated optimization of a course of actions in a stochastic environment"
author: "Thomas Vicente"
output: pdf_document
fontsize: 12pt
header-includes: \usepackage{bm}
                 \usepackage{bbm}
                 \usepackage{algorithmic}
                 \usepackage{algorithm}
---

#General principle

\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
\begin{algorithm*}
\caption{General version}
\begin{algorithmic}
\STATE$\textbf{initialize:}$
\IND$f() \text{: regressor function}$
\IND$x_t \text{: current stochastic environment}$
\IND$\Omega_t(x_t) \text{: set of possible actions}$
\IND$s(x_t) \text{: signal activated or not by the stochastic environment, initialized to 0}$
\IND$W^*=W_0\text{: random weights for the regressor}$
\IND$X\text{: empty training feature matrix}$
\IND$y\text{: empty training target vector}$
\IND$A\text{: empty feature matrix for simulated actions}$
\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
\IND$\Delta = 0 \text{: number of actions}$
\STATE$\textbf{while } \text{the learning process is on:}$
\IND$\text{record }x_t$
\IND$A \text{ = empty matrix}$
\IND$\textbf{for } a' \text{ simulated action in } \Omega_t(x_t)$
\IND[2]$A.append(a')$
\IND$a^*=\arg\max_{a' \in A} f(W^*,A)$
\IND$X.append(a^*)$
\IND$\textbf{if } s(a^*) \text{ != } 0 \text{ :}$
\IND[2]$\textbf{for } a \text{ in } \Delta \text{:}$
\IND[3]$value = \gamma^{a-1}s(a^*)$
\IND[3]$y.append(value)$
\IND[3]$W^* := \arg\min_{W}(l(f(W,X),y))$
\IND[3]$\Delta = 0$
\IND$\textbf{else} \text{:}$
\IND[2]$\Delta \text{ += } 1$
\IND[2]$\text{continue}$
\end{algorithmic}
\end{algorithm*}

The algorithm assumes two underlying function:

- The minimization process is a tedious part of the algorithm. When dealing with unstructured features, we might want to use a neural network-type regressor. In that case, $W$, the set of the hidden layers' weights is initialized with the values of the previous iterations. The optimal $W^*$ should converge as $X$ grows if there are patterns in the stochastic processes.

- The signal function assumes the existence of one or multiple sensors, or the manual assignement of a value. It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.

###Remark on the function of the regressor's role
The regressor's task is not to predict accurately, even though it is closely related, but rather to give the "real value" of an action.

#Analysis of why RL+NN converge, why it works
It needs to have a good balance between gains and pains

#Application

###Remark on the function of the regressor's role
Two intermediary movements can ultimately lead to both winning and losing. Such states will get a non-tendencial value as $X$ grows, which is a good thing.

##Analysis of game
Do some graph based on data.txt's target variable, observe time effects, wins evolution

### 2016-05-25 imgresize.py

@@ -3,16 +3,36 @@ import os
 os.chdir('/Users/Thomas/git/thesis/images6')
 dim = 3
 
+def normalize(arr):
+    arr = arr.astype('float')
+    for i in range(3):
+        minval = arr[...,i].min()
+        maxval = arr[...,i].max()
+        if minval != maxval:
+            arr[...,i] -= minval
+            arr[...,i] *= (255.0/(maxval-minval))
+    return arr
+
 for i in os.listdir(os.getcwd()):
 	if i.endswith('png'):
 		if i.startswith('X') or i.startswith('O') or i.startswith('N'):
 			img = Image.open(i).resize((dim,dim), Image.ANTIALIAS).convert('LA')
+			bright = ImageEnhance.Brightness(img)
+			img = bright.enhance(1.3)
+		#elif i.startswith('N'):
+		#	img = Image.open(i).resize((dim,dim), Image.ANTIALIAS).convert('LA')
+		#	bright = ImageEnhance.Brightness(img)
+		#	img = bright.enhance(1.55)
 		elif i.startswith('g'):
 			img = Image.open(i).resize((dim*3,dim*3), Image.ANTIALIAS).convert('LA')
-		bright = ImageEnhance.Brightness(img)
-		img = bright.enhance(1.5)
 		contrast = ImageEnhance.Contrast(img)
 		img = contrast.enhance(3)
+
 		img.save('smalls/'+i)
+
+		#img = Image.open('smalls/'+i).convert('RGBA')
+		#img = np.array(img)
+		#img = Image.fromarray(normalize(img).astype('uint8'),'RGBA')
+		#img.save('smalls/'+i)
 	else:
 		continue

### 2016-05-25 tic19.py

@@ -1,33 +1,8 @@
-
-
-
-
-
-
-
 #TO DO:
     #IMG DISTORT: http://stackoverflow.com/questions/14177744/how-does-perspective-transformation-work-in-pil
     #convolutional
-    #extra: not just where it plays but heatmap according to predict values
-
-
-
-
-
-
-    #NORMALIZE IMAGE: http://stackoverflow.com/questions/1735025/how-to-normalize-a-numpy-array-to-within-a-certain-range
-    #maybe just before training
-
-
-
-
-
-
-
-
-
-#best: danielnouri.org/notes/2014/12/17/using-convolutional-neural-nets-to-detect-facial-keypoints-tutorial/
-#One advantage is that training converges much faster; maybe four times faster in this case. The second advantage is that it also helps get better generalization; pre-training acts as a regularizer. 
+    #best: danielnouri.org/notes/2014/12/17/using-convolutional-neural-nets-to-detect-facial-keypoints-tutorial/
+    #One advantage is that training converges much faster; maybe four times faster in this case. The second advantage is that it also helps get better generalization; pre-training acts as a regularizer. 
 
 from __future__ import division
 from random import sample
@@ -43,7 +18,7 @@ import sys
 import scipy
 from scipy import misc
 from sknn.mlp import Regressor, Layer, Convolution
-from PIL import Image, ImageEnhance, ImageDraw
+from PIL import Image, ImageEnhance, ImageDraw, ImageOps
 from lasagne import layers as lasagne, nonlinearities as nl
 import glob
 
@@ -56,7 +31,7 @@ maxi = 0
 pred = 0
 who = 'none'
 dim = 3
-selfplays = 1
+selfplays = 100
 for i in os.listdir(smalls):
     if i.startswith('X'):
         Xs.append(scipy.misc.imread(i, flatten=True))
@@ -144,7 +119,7 @@ class Learner:
                 learning_rule='rmsprop',
                 f_stable=0.0001,
                 verbose=True,
-                batch_size=100,
+                batch_size=200,
                 n_stable=10)
 
         return s.net.fit(X, y)
@@ -310,31 +285,32 @@ class Game:
         if s.state[j,i] == 0:
             #import and treat one image per move
             os.chdir('/Users/Thomas/git/thesis/newgames')
-            newest = max(glob.iglob('*.[Pp][Nn][Gg]'), key=os.path.getctime)
[…]

### 2016-05-28 paper.Rmd
te

@@
 #General principle
+Explain how it works in sentences
 
@@
 \begin{algorithmic}
-\STATE$\textbf{initialize:}$
-\IND$f() \text{: regressor function}$
-\IND$x_t \text{: current stochastic environment}$
-\IND$\Omega_t(x_t) \text{: set of possible actions}$
-\IND$s(x_t) \text{: signal activated or not by the stochastic environment, initialized to 0}$
-\IND$W^*=W_0\text{: random weights for the regressor}$
-\IND$X\text{: empty training feature matrix}$
-\IND$y\text{: empty training target vector}$
-\IND$A\text{: empty feature matrix for simulated actions}$
+\STATE$\text{Initialized:}$
+\IND$X\text{: empty }m\text{-columns feature matrix}$
+\IND$y\text{: empty target vector}$
+\IND$X'\text{: empty }m\text{-columns simulated states matrix}$
+\IND$w^*=w_0\text{: random weights for the regressor}$
 \IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
 \IND$\Delta = 0 \text{: number of actions}$
-\STATE$\textbf{while } \text{the learning process is on:}$
+\STATE$\text{Other variables:}$
+\IND$f() \text{: regressor function}$
+\IND$m \text{: number of features}$
+\IND$p_t(x^*) \text{: stochastic process, function of the last existing chosen state}$
+\IND$x_t(p_t) \in {\rm I\!R^m} \text{: current state vector}$
+\IND$\Omega_t(x_t) \text{: set of possible actions}$
+\IND$s(x_t) \in {\rm I\!R} \text{: signal value triggered by the modified state}$
+\STATE$\textbf{WHILE } \text{the learning process is on:}$
 \IND$\text{record }x_t$
-\IND$A \text{ = empty matrix}$
-\IND$\textbf{for } a' \text{ simulated action in } \Omega_t(x_t)$
-\IND[2]$A.append(a')$
-\IND$a^*=\arg\max_{a' \in A} f(W^*,A)$
-\IND$X.append(a^*)$
-\IND$\textbf{if } s(a^*) \text{ != } 0 \text{ :}$
-\IND[2]$\textbf{for } a \text{ in } \Delta \text{:}$
-\IND[3]$value = \gamma^{a-1}s(a^*)$
-\IND[3]$y.append(value)$
-\IND[3]$W^* := \arg\min_{W}(l(f(W,X),y))$
-\IND[3]$\Delta = 0$
-\IND$\textbf{else} \text{:}$
+\IND$X' := \text{ empty matrix}$
+\IND$\textbf{FOR } x' \text{ in } \Omega_t(x_t)\text{:}$
+\IND[2]$\text{append } x' \text{ to } X'$
+\IND$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
+\IND$\text{append } x^* \text{ to } X$
+\IND$\textbf{IF } s(x^*) \text{ != } 0 \text{ :}$
+\IND[2]$\textbf{FOR } \delta \text{ in } 1\text{:}\Delta \text{:}$
+\IND[3]$value = \gamma^{\delta-1}s(x^*)$
+\IND[3]$\text{append } value \text{ to } y$
+\IND[2]$w^* := \arg\min_{w}(l(f(w,X),y))$
+\IND[2]$\Delta := 0$
+\IND$\textbf{ELSE} \text{:}$
 \IND[2]$\Delta \text{ += } 1$
-\IND[2]$\text{continue}$
 \end{algorithmic}
@@
 
-The algorithm assumes two underlying function:
+Explanation of some elements:
+
+- $\Omega$ can be set deterministically.
+In the context of a game, there is a well determined space of actions.
+It also can be infinite.
+A mobile robot can explore the real 3D world in almost infinite ways.
+
+- The signal value function assumes the existence of one or multiple sensors, or the manual assignement of a value.
+It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.
 
@@
 The optimal $W^*$ should converge as $X$ grows if there are patterns in the stochastic processes.
-
-- The signal function assumes the existence of one or multiple sensors, or the manual assignement of a value.
-It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.
 

### 2016-05-28 autocontrast.py

@@ -1,0 +1,10 @@
+from PIL import Image, ImageEnhance, ImageOps
+import os
+import numpy as np
+os.chdir('/Users/Thomas/git/thesis/normalizetests')
+dim = 3
+for i in os.listdir(os.getcwd()):
+	if i.endswith('png') and i.startswith('g'): 
+		img = Image.open(i)
+		img = ImageOps.autocontrast(img,ignore=range(0,135)+range(230,256))
+		img.save('norm'+i)

### 2016-05-28 paper.Rmd
df

@@
 
+#Introduction
+First sentence: ~philosophical. A complete exercise would be to build an autonomous entity that is able to take long term oriented decisions based on information coming from the phisical reality.
+
+This entity needs two abilities: learning from its own actions and recognising the current environment's patterns.
+The first ability is typically made possible by a Reinforcement Learning (RL) mechanism.
+The second ability is done by regressors or classifiers powerful enough to adapt to the entity's sensorial features.
+In particular, if the entity is equipped by a visual sensor, neural networks might be a good idea.
+
+RL emerges easily from other Machine Learning techniques as the training is done by acting.
+The learner evolves in particular environment and records what actions lead to a maximal gain.
+In most cases the goal is to maximise a long term gain by taing a correct course of actions.
+These policies can be more or less explicit.
+An RL learner is particularly dependent on a state signal that succeeds in retaining all relevant information ; this state is said to be Markov.
+For example, a checkers position (the current configuration of all the pieces on the board) would serve as a Markov state because it summarizes everything important about the complete sequence of positions that led to it.
+Much of the information about the sequence is lost, but all that really matters for the future of the game is retained.
+
+DO I DO ON OR OFF POLICY? Policy related to Markov decision process
+
+Examples of regressors/classifier "recognising the real world"
+Using an approximator has two benefits: ability to provide an evaluation of an infinite set of states and recognise the underlying states (equivalent for the purpose of a particular task) when the signal is imperfect, which is our case with our study.
+
+Tic tac toe, a classical RL application, is also a good example for the principle studied. I present an application of the approchaed principle in the second part.
+
+#Litterature review
+
+The closest algorithm from our principle is in 9.4 in Sutton (2012).
+But the proposed principle differs from it.
+First, we consider the possibility of same underlying states with imperfect signals (in a handwritten game, this is due to the differences in handwritten letters every time a human plays).
+Second, we do not use Temporal differences but a discount method.
+
+Deep Q networks (DQN) are similar to the proposed principle as they store long term data.
+This is the so called "Experience replay": when training the network, random minibatches from the replay memory are used instead of the most recent transition.
+Such efficient implementation makes the DQN very adaptive to different tasks, as demonstrated for the learning of multiple arcade games while using the same network architecture.
+
 #General principle
 Explain how it works in sentences
+
+Alternative discount
+
+Table of all possible values from discounted equation (about 12 values I guess)
+
+quick review of different regressors
+
+In the algorithm description, we try to get a maximum of generality and thus use the term "regressor".
+
+Seems that "discount post signal" is the particularity of model.
+Give empirical insights why it works/is efficient.
+Good point: each step takes a value, not only final ones (can represent with a colored gradient picture)
 
@@
 \IND[2]$\text{append } x' \text{ to } X'$
-\IND$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
+\IND$\text{With probability } 1-\epsilon$
+\IND[2]$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
+\IND$\text{With probability } \epsilon$
+\IND[2]$x^*:=\text{random sample }x' \in X'$
 \IND$\text{append } x^* \text{ to } X$
@@
 It needs to have a good balance between gains and pains
+Find some equations, ask Gabor then
+
+#Cool remark
+sometimes the triggering a new signal can take time. But robot can still look for best discounted s(x^*). Wecan consider that the robot "discovers" new objectives in that context. True in general RL
 
 #Application
+
+Explain game
+"Long term" (3-5 actions) strategy matters. Discount acts here
+The environment contains rivalry. Another, more classical, robot.
+
+###Talk about tweaks
+generation of images
+other guy for selfplay...
+rotations to go faster than Other
+learn not after each game
 
@@
 
+why chose NN. explain how it works. choice of architecture
+
 ##Analysis of game
 Do some graph based on data.txt's target variable, observe time effects, wins evolution
+Proof of convergence
+
+The following plots show the evolution the robot's performance during the learning phase.
+Recall that the robot updates its parameters every 100 (MIGHT CORRECT) games.
+The green line corresponds to the proportion of won games for the last 100 games The red line corresponds to the proportion of lost games for the last 100 games
+
+```{r,echo=F}
+#show in grid, different epsilons for 1000 games, 10000 also?
+setwd('/Users/Thomas/git/thesis/reports/')
+
+plotsmthg = function(what, nbg, random, colour,yrange){
+  games = seq(nbg/10,nbg,nbg/10)
+  d = read.csv(paste0(nbg,'games_epsilon',random,'.txt'), header=F)
+  plot(games, d[,what], type='l', main = paste0('Randomness of the oponent=',random), xlab='Number of games', ylab='', bty='n', col=colour,ylim=yrange)
+}
+
+plotsmthg(1,100,0.99,'green',range(0,1))
+par(new=T)
+plotsmthg(2,100,0.99,'red',range(0,1))
+par(new=F)
+plotsmthg(3,100,0.99,'black',range(3,4))
+```
+
+The last plot shows the average number of moves before winning for the last 100 games.
+
+#Notice after 100 games, able to play "against" image if move easy, like last action
+
+#Results/other applications
+
+#References
+http://people.inf.elte.hu/lorincz/Files/RL_2006/SuttonBook.pdf
+
+http://www0.cs.ucl.ac.uk/staff/D.Silver/web/Teaching_files/FA.pdf

### 2016-05-28 tic19.py

@@ -6,7 +6,9 @@
 
 from __future__ import division
 from random import sample
+import datetime
 import random
+import csv
 import numpy as np
 from numpy import diag
 import cPickle
@@ -31,7 +33,7 @@ maxi = 0
 pred = 0
 who = 'none'
 dim = 3
-selfplays = 100
+selfplays = 1000
 for i in os.listdir(smalls):
     if i.startswith('X'):
         Xs.append(scipy.misc.imread(i, flatten=True))
@@ -83,7 +85,7 @@ class Learner:
         if player == 1:
             if alpha == None: s.alpha = 0.8
             else: s.alpha = alpha
-            if epsilon == None: s.epsilon = 0.5
+            if epsilon == None: s.epsilon = 0.99
             else: s.epsilon = epsilon
         else: 
             if alpha == None: s.alpha = 0.8
@@ -365,6 +367,7 @@ class Selfplay:
         s.i = 0
         s.games = 0
         s.wins = 0
+        s.losses = 0
         s.wining = []
         s.valuesave = dict()
 
@@ -373,6 +376,7 @@ class Selfplay:
 
         s.y_game = np.empty((0))
         s.moves = 1
+        s.winsteps = []
         s.gamma = 0.7
 
     def reset(s):
@@ -455,7 +459,8 @@ class Selfplay:
                 s.y_game = np.repeat(s.y_game, 4, axis=0) #duplicate each row 4 times
                 s.y2 = np.concatenate((s.y2, s.y_game), axis=0)
                 s.y_game = np.empty((0))
-                s.moves = 1
+                if s.state.full() or s.state.won(1):
+                    s.moves = 1
             else:
                 s.moves += 1
 
@@ -481,18 +486,38 @@ class Selfplay:
 
                 if s.state.won(2):
                     s.wins += 1
+                    s.winsteps.append(s.moves)
+                    s.moves = 1
+                if s.state.won(1):
+                    s.losses += 1
 
             #learn from updated matrix every selfplays/10 games
             if s.games %(selfplays/10) == 0 or s.games == selfplays:
                 if s.state.full() or s.state.won(1) or s.state.won(2):
-                    print 'wins proportion: ', s.wins/s.games
+
+                    propwin = s.wins/s.games
+                    proploss = s.losses/s.games
+                    meanwinsteps = reduce(lambda x, y: x + y, s.winsteps)/len(s.winsteps)
+                    obs = [propwin, proploss, meanwinsteps]
+                    print 'obs ', obs
[…]

### 2016-05-30 paper.Rmd
yep

--- 2016-05-28 paper.Rmd
@@
          \usepackage{algorithm}
+
+abstract: "I propose a principle allowing an artificial learning entity to find a right course of action possessing (like we all do) only an imperfect set of information on the current state of the world.
+I try to desmontrate the practicality and the generality of such a principle."
+
+keywords: {reinforcement learning, neural networks, artificial intelligence}
 ---
 
-#General principle
-Explain how it works in sentences
+#1 Introduction
+A complete exercise would be to build an autonomous entity that is able to take long term oriented decisions based on information coming from the phisical reality.
+
+This entity needs two abilities: learning from its own actions and recognising the current state's patterns.
+The first ability is typically made possible by a Reinforcement Learning (RL) mechanism.
+The second ability is done by regressors or classifiers powerful enough to adapt to the entity's sensorial features.
+In particular, if the entity is equipped by a visual sensor, neural networks might be a good idea.
+
+##1.1 Learning from its own actions
+
+RL emerges easily from other Machine Learning techniques as the training is done by acting.
+It uses training information that *evaluates* the actions taken rather than *instructs* by giving correct actions are faced repeatedly with a choice among n different options, or actions.
+Such setting has its roots in the *n-armed bandit problem* where after each choice you receive a numerical reward chosen from a stationary probability distribution that depends on the selected action.
+The objective is to maximize the expected total reward over all the episodes.
+The learner evolves in a particular state and records what actions lead to a maximal gain.
+In most cases the goal is to maximise a long term gain by taing a correct course of actions.
+These policies can be more or less explicit and the proposed algorithm would qualified as *off-policy*.
+
+An RL learner is thus particularly dependent on a state signal that succeeds in retaining all relevant information, a *Markov* state.
+For example, a checkers position (the current configuration of all the pieces on the board) would serve as a Markov state because it summarizes everything important about the complete sequence of positions that led to it.
+Much of the information about the sequence is lost, but all that really matters for the future of the game is retained.
+
+A reinforcement learning task that satisfies the Markov property is called a Markov Decision Process (MDP).
+If the state and action spaces are finite, it is possible the MDP framework notably formalizes the probability that a state occurs given a previous state an action, and the probability of reward after the new state occured.
+
+##1.2 Recognising real world situations
+
+Equipping a reinforcement learner with an approximator function has two benefits: the ability to provide an evaluation of an infinite set of states and recognise the a finite set of states having an infinite set of possible representations.
+In the example below, as humans we are able to recognise the same underlying situation and take action: if I learner how to make spaghettis at home, I am able to reproduce it at my friend's place.
+In a similar manner, we would like our learner to overcome its imperfect sensors to learn how to behave in the long term to maximize its gain.
+This is made possible notably by regressor and classifier functions that take particularly well advantage of naturally generated data such as images and sounds.
+
+A modified version of the classical tic-tac-toerobot will provide a good example for the principle we seek to study.
+
+TWO IMAGES, caption: after enough games, the learner should be able to play the same way after vizualising either of these images
+
+#2 Litterature review
+
+The proposed algorithm allows to optimize a course of actions in finite and infinite spaces. But it has especially been thought to act based on imperfectly sensed states as described above.
+
+A this is where it differs for example from the State-Action-Reward-State-Action aglorithm (SARSA).
+It uses approximations to update the value of a whole set of actions, a policy.
+SARSA is well suited when the value of each action can be monitored.
+It notably update the parameters of the approximation function after each movement thanks to a change in value from $t-1$ to $t$, known as a temporal difference.
+The update incldues a gradient term, compute from the difference between the state's value approximation and the real state's value.
+
+The Deep Q networks (DQN) are close to what I propose.
+This consists in an off-policy algorithm using neural network approximations of the states' value to decide on an action.
+And as the learner makes better moves, the neural network is fed with more useful information.
+And thanks to the so called "Experience replay", when training the network, random minibatches from the replay memory are used instead of the most recent transition.
+Such efficient implementation makes the DQN very adaptive to different tasks, as demonstrated for the learning of multiple arcade games while using the same network architecture.
+
+Hopefully the proposed algorithm offers even more flexibility than the evoked algorithms.
+That would be thanks to removing the assumption that the state's values are permanently monitored during the training phase.
+This is possible by the implementation of a *signal function*.
+Its role is to detect a preset objective signal thanks to a deterministic function or a well tuned sensor.
+The aim is notably to allow the algorithm to be practial when evaluating non-idealized states - unlike video game screenshots - involving stochastic perturbations - such as real world pictures or videos.
+
+#3 Proposed principle
+
+##3.1 Description
+To describe the proposed principle, we are going to replicate chronologically an episode of the algorithm in action.
+
+We call the learning algorithm the *learner*. The learner's aim is to automate a course of actions $x^*$ observing at $t$ a stochastic state $x_t$. Notice that $x^*$ is also equivalent to a new state
+
+At each episode, the learner is facing different choices of actions.
+The space of possible actions is called $\Omega$.
+It can be a discrete or continuous space.
+It can be set deterministically ; for instance, in the context of a game, there is a well determined space of actions that a human can tune in advance.
+It also can be infinite ; for instance, a mobile robot can explore the real 3D world in almost infinite ways.
+
+The choice, with probability $1-\epsilon$, which corresponds to a non-exploratory move, is called $x^*$, the maximum over a set of predictions.
+These predictions correspond to a regressor function that we describe below.
+These predictions are made over $X'$, a set of simulated actions belonging to $\Omega$.
+$x^*$ is then added to a training set, $X$.
+
+Now, we are assuming that the learner posesses at least one sensor allowing it to get an objective signal.
+This signal, $s(x^*)$, is a function of the state generated by the learner's last move.
+It is conditional on the current state and takes positive value if a *gain* is sensed, and negative value if a *loss* is sensed.
+At least one value should be potentially triggered.
+A typical example for $[gain;loss]$ would be respectively $[1;-1]$.
+They should be set so they are proportional to the learner's closeness to its objective.
+
+While this signal is off, the learner counts the number of actions since the signal was last triggered.
+If this signal is triggered, the current episode is over and each action since the last signal was triggered is receiving a value given by the function $d(\gamma, \Delta, \delta, s(x^*))$.
+$\gamma$ is a discount rate, $\Delta$ the number of actions in the episode and $\delta$ the index of an action in the episode (the first action in the episode has $\delta=1$).
+So the closer an action was from the triggered signal, the larger the magnitude of its value will be.
+Each value corresponding to all the last actions are appended to the training target vector $y$.
+I realize this is only a particular way to discount these values, and different ways might be considered for different applications.
+
+We then get a first training sample on which are going to apply a regressor function $f()$.
+Its aim is to minimize the loss $l(f(w,X),y)$.
+The parameter set corresponding to the minimal loss is $w^*$.
+It is going to be used in the next episode in order to predict the value of simulated actions
+
+Here is a formal version of the algorithm:
+
+\newpage
 
@@
 \begin{algorithm*}
-\caption{General version}
+\caption{}
 \begin{algorithmic}
@@
 \IND$\Omega_t(x_t) \text{: set of possible actions}$
-\IND$s(x_t) \in {\rm I\!R} \text{: signal value triggered by the modified state}$
+\IND$\epsilon \text{: randomness rate}$
+\IND$s(x^*) \in {\rm I\!R} \text{: signal value triggered by the modified state}$
 \STATE$\textbf{WHILE } \text{the learning process is on:}$
@@
 \IND[2]$\textbf{FOR } \delta \text{ in } 1\text{:}\Delta \text{:}$
-\IND[3]$value = \gamma^{\delta-1}s(x^*)$
+\IND[3]$value = d(\gamma, \Delta, \delta, s(x^*))$
 \IND[3]$\text{append } value \text{ to } y$
@@
 
-Explanation of some elements:
+\newpage
 
-- $\Omega$ can be set deterministically.
-In the context of a game, there is a well determined space of actions.
-It also can be infinite.
-A mobile robot can explore the real 3D world in almost infinite ways.
+##3.2 Remarks
 
-- The signal value function assumes the existence of one or multiple sensors, or the manual assignement of a value.
-It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.
+###3.2.1 Regressor function
 
-- The minimization process is a tedious part of the algorithm.
-When dealing with unstructured features, we might want to use a neural network-type regressor.
-In that case, $W$, the set of the hidden layers' weights is initialized with the values of the previous iterations.
-The optimal $W^*$ should converge as $X$ grows if there are patterns in the stochastic processes.
+Depending of the sensor the learner has at his disposal, some regressor functions can be more or less suited.
 
-###Remark on the function of the regressor's role
-The regressor's task is not to predict accurately, even though it is closely related, but rather to give the "real value" of an action.
+If the possible states form a linear function of the target value, a linear regressor function can be applied.
+But the states can be strongly affected by stochastic processes.
+For structured data, non-parametric tree-based methods might be considers.
+Our typical case is when this happens with unstructured data, such as images, sounds and texts.
+Then, neural networks might be considered.
 
-#Analysis of why RL+NN converge, why it works
-It needs to have a good balance between gains and pains
+This last case is typically mimicking how human would act if faced with a particular situation.
+The challenge is here to recognise equivalent underlying states thanks to pattern recognition.
+And considering the algorithm described above, $w$, would be the set of a neural network's hidden layers.
 
-#Application
+One interesting aspect is the role of the regressor function.
+Contrarily to typical applications, the functions's task is utimately not to predict an action's value with maximum accuracy, even though it is often a dependency, but rather to give the "real value" of an action.
+Imagine for instance a game where two intermediary states might lead to both winning and losing.
+Such states will get a non-tendencial value from the regressor function, which is a good thing.
 
-###Remark on the function of the regressor's role
-Two intermediary movements can ultimately lead to both winning and losing. Such states will get a non-tendencial value as $X$ grows, which is a good thing.
+###3.2.2 Signal function and discounted values
 
-##Analysis of game
-Do some graph based on data.txt's target variable, observe time effects, wins evolution
+The signal function allows for minimal assumptions on the basic abilities of the learner.
+While most approximation-based reinforcement learning algorithms suppose the permanent presence of a value monitoring function, here we only assume the presence of a discrete, time-independent sensor triggering one or more values.
+
+The use of this simple sensor is made possible by the use of discounted
+values.
+Even if not every action is monitored during an episode, each step still takes a value at the end of the episodes.
+In that process, tuning correctly the $\gamma$ parameter is important.
+In the following figure, we represent two sets of actions that to respectively a gain, and a loss.
+
+for gain and loss, display same gradient: from red to green
+SHOW IMAGE TREE WITH DIFFERENT DISCOUNT FACTORS
+
+We see that with a higher discount factor, a stronger emphasis is put on the final result of the set of actions.
+In some cases, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.
+
+With this function, triggering a new signal can take time.
+But the learner can still look for best discounted value according to the current state.
+We then benefit from something common to RL algorithms: the use of intermediary objectives.
+
+#4 Asymptotic behaviour of the algorithm
+
+include idea RL+regressor (Gabor)
+
+##4.1 Convergence for equivalent underlying states
+
+The convergence of the algorithm's evaluations is closely linked to the performance of the regressor function at extracting underlying patterns between states and values, assuming however that this relationship exists.
+
+Let's consider a finite set of underlying states.
+If the conditions are favorable, we expect overtime, as the training sample is fed after each episode, a better evaluation of the different states and, asymptotically, underlying equivalent states will get the same value.
+Under another perspective, we expect $w^*$ to converge as $X$ grows.
+
+An interesting point is the expected non-linearity of this process.
+Indeed, as long as the learner did not discover all gain-yielding states, the regressor's parameters do not permanently stabilize.
+Once all states have been uncovered, this convergence happens.
+
+If the set of underlying states is infinite, the regressor's parameters can stabilize only if there exist patterns between the states and the value.
+An example of this situation is a vision-equipped learner whose task is to catch the same ball in different environments.
+
+##4.2 Long term gain maximization
+
+Evaluating well the value of equivalent states is only a first step to obtain a learner that is useful in practice.
+To behave in a long term-maximizing way, the algorithm needs to have a relatively balanced experience between gains and losses.
+
+First, tuning $\gamma$ in a proper way will help to allow the regressor to evaluate states well faster, via the learner's actions. If possible, a good way to tune this parameter is to make such that
+$$d(\gamma, \Delta, \delta, s_{gain}) \simeq d(\gamma, \Delta, \delta, s_{loss})$$
+for the $\delta$th intermediary action that led to a gain in one episode but also to a loss in another episode.
+
+Second, tuning $\epsilon$ well will insure that the learner explores potentially good actions fast enough. A number of episodes-dependent $\epsilon$ might be desirable in that regard.
+
+#5 Implementaion for a "real-world" tic-tac-toe game
+
+The tic-tac-toe game is a game with initially nine positions organised as a three by three square.
+After the first player played the first move, the second player is left with eight available moves.
+This is until one player won or once the whole board is covered, yielding a draw.
+To win, a player must align his moves on the same line or on the same column, or by filling one of the board's diagonal.
+
+The tic-tac-toe game is a RL classical application, but this time the learner has to play according to his "vision".
+Notably, when playing against a human, the information available to the robot corresponds to an image of the type:
+PICTURES OF VISION
+caption: pictures of the same underlying state
+This is thus a "stochastic representation" from a finite set of states. The stochastic processes affecting this representation are notably the varying handwrittings and the environment's brightness.
+
+After a sufficient number of games, the robot is able to effectively assign values to the potential actions it might take:
+ADD 150*150 PICTURE WITH PROBAS ON TOP, FOR 100, 1000, 5000 selfplays
+REVALUE SO LOSS = -1, DRAW=0
+
+##5.1 Particularities of the implementation
+
+The stochastic state $x_t$ is resulting from another player's move represented on an image as seen above.
+In terms of strategy optimization, here the learner would have to think ahead.
+The states here are the fruit of *rival* actions, either from a human player or from another robot.
+And in this implmentation, the learner plays after the rival, so it can win either in three or four moves.
+
+The signal function is determincally set by a human.
+It triggers the values $1$ if the learner wins, $-0.1$ if there is a draw and $-1$ if it loses or if it does an illegal move.
+Notice that an illegal move corresponds only to playing on top of the other player.
+The learner here knows where it last played.
+
+In this implementation, the discount function $d(\gamma, \Delta, \delta, s(x^*))$ has the form
+$$\gamma^{\Delta-\delta}s(x^*)$$
+More precisely, I picked $\gamma=0.3$ as it yields the same value for an early move.
+The discounted value for moves leading to a win thus are: 1, 0.3, 0.09 and 0.081.
+For a draw: -0.1, -0.01, -0.001 and -0.0001.
+For a loss: -1, -0.3, -0.09 and -0.081.
+For an illegal move, only the signal value is given to the last move.
+
+The chosen regressor is an artificial neural network.
+I chose this type of regressor as I believe it supports the complex relationship between the pixels representing a state and the value of this state.
+A sufficiently complex structure is thus able to handle the very large number of ways to represent a game combined with the many game combinations.
+In this implementation, the network is has the following form:
+PICTURE NN
+
+##5.2 Necessary adjustments
+
+A few adjustments were to be made.
+First, notice that the learning phase is done against a Q-learning robot.
+After each of the players' move, an image representing the current state of the board is generated.
+This is image is then reduced and appended to the training set.
+SHOW LARGE IMAGE -> SMALL IMAGE transition
+
+Second, our learner had the disadvantage of receiving imperfect information about the game compared to the other robot, knowing exactly where the moves are made.
+To compensate for this, our learner after each move rotates the game three times and appends the learned values accordingly.
+
+Third, as we do not have an objective sensor at our disposal, the signal recieved by the learner is the same as the signal coming from the oponent robot.
+It is an objective information indicating the final state of game when it happens.
+
+Third, the learning process, via neural networks, does not happen after each game but at regular intervals.
+This is purely for computational constraints.
+However, this does not affect in fine the performance of the learner, which is able to win up to 90% of the last phases' games.
+
+##5.3 Performance analysis
+
+This section aims at assessing that the proposed algorithm leads to winning strategies.
+
+The following plots show the evolution the learner's performance during the learning phase.
+We also study how the proportions of wins and losses (and draws by deduction) vary according to the levels of $\epsilon$ for each type of robot.
+Recall that the robot updates its parameters every 100 games, and can potentially improve at this moment.
+
+The green line corresponds to the proportion of won games for the last 100 games The red line corresponds to the proportion of lost games for the last 100 games.
+
+```{r,echo=F}
+#show in grid, different epsilons for 1000 games, 10000 also?
+setwd('/Users/Thomas/git/thesis/reports/')
+
+plotsmthg <- function(what, nbg, random, colour, yrange){
+  games <- seq(nbg/10,nbg,nbg/10)
+  d <- read.csv(paste0(nbg,'games_epsilon',random,'.txt'), header=F)
+  plot(games, d[,what], type='l', main = paste0('Randomness of the oponent=',random), xlab='Number of games', ylab='', bty='n', col=colour,ylim=yrange)
+}
+
+plotsmthg(1,100,0.99,'green',range(0,1))
+par(new=T)
+plotsmthg(2,100,0.99,'red',range(0,1))
+
+#plotsmthg(3,100,0.99,'black',range(3,4))
+```
+
+We see that a more eploratory learner improves its winning rate faster but is then relatively limited towards the last phases.
+
+A more exploratory oponent helps the learner to win and thus discover what are the winning strategies.
+However, making the oponent too exploratory would limit the performance of the learner when playing against a trained human.
+
+INCLUDE ADAPTATIVE LEARNING RATE?
+
+Notice that a learner trained against a random oponent is able, after 100 games, to play easy moves, typically the last move to win, against a human player.
+
+The last plot shows the average number of moves before winning for the last 100 games.
+
+PLOT AVG MOVES
+
+The learner can only win in three or four moves. Against a random player, we that the learner has an average number of moves to win closer to three with more games.
+
+#6 Conclusion
+
+The proposed algorithm behaved as expected with its tic-tac-toe implementation.
+More work has to be done concerning the proof of convergence for the regressor function and the proof that asymptotically the behaviour of the learner is optimal.
+
+Now, we can think of more useful applications than this overkill tic-tac-toe player.
+For instance, a learner of this type, installed on augmented reality glasses could help people achieving precise tasks in a more optimal way.
+It could also be useful in online marketing to automatically identify set of actions leading to better sales, such as the sequence of clicks of customers.
+We also can imagine learning (thanks to vision) the way natural organisms evolve in their environment to build very adaptative machines.
+
+#8 References
+Sutton and Barto, *Reinforcement learning: An introduction*, Vol.
+2.
+No.
+1, MIT press, 2012, available at [people.inf.elte.hu/lorincz/Files/RL_2006/SuttonBook.pdf](people.inf.elte.hu/lorincz/Files/RL_2006/SuttonBook.pdf)
+
+http://www0.cs.ucl.ac.uk/staff/D.Silver/web/Teaching_files/FA.pdf
+
+https://www.cs.toronto.edu/~vmnih/docs/dqn.pdf
+
+http://gaips.inesc-id.pt/~fmelo/pub/melo08icml.pdf
+
+*Human-level control through deep reinforcement learning* http://www.readcube.com/articles/10.1038/nature14236?shared_access_token=Lo_2hFdW4MuqEcF3CVBZm9RgN0jAjWel9jnR3ZoTv0P5kedCCNjz3FJ2FhQCgXkApOr3ZSsJAldp-tw3IWgTseRnLpAc9xQq-vTA2Z5Ji9lg16_WvCy4SaOgpK5XXA6ecqo8d8J7l4EJsdjwai53GqKt-7JuioG0r3iV67MQIro74l6IxvmcVNKBgOwiMGi8U0izJStLpmQp6Vmi_8Lw_A%3D%3D

### 2016-05-31 paper.Rmd
fdsfdf
dfdf

@@
 In most cases the goal is to maximise a long term gain by taing a correct course of actions.
-These policies can be more or less explicit and the proposed algorithm would qualified as *off-policy*.
+These policies can be more or less explicit.
+I our case, the proposed algorithm would be qualified as *off-policy*.
 
@@
 
-##1.2 Recognising real world situations
+##1.2 Pattern recognition
 
@@
 
-#3 Proposed principle
+#3 Proposed algorithm
 
@@
 \begin{algorithmic}
-\STATE$\text{Initialized:}$
+\STATE$\text{Initialized variables:}$
 \IND$X\text{: empty }m\text{-columns feature matrix}$
 \IND$y\text{: empty target vector}$
-\IND$X'\text{: empty }m\text{-columns simulated states matrix}$
 \IND$w^*=w_0\text{: random weights for the regressor}$
@@
 \IND$p_t(x^*) \text{: stochastic process, function of the last existing chosen state}$
+\IND$X'\text{: }m\text{-columns simulated states matrix}$
 \IND$x_t(p_t) \in {\rm I\!R^m} \text{: current state vector}$
@@
 \IND$\epsilon \text{: randomness rate}$
-\IND$s(x^*) \in {\rm I\!R} \text{: signal value triggered by the modified state}$
+\IND$s(x^*) \in {\rm I\!R} \text{: signal value potentially triggered by the modified state}$
 \STATE$\textbf{WHILE } \text{the learning process is on:}$
-\IND$\text{record }x_t$
-\IND$X' := \text{ empty matrix}$
+\IND$\text{record }x_t(p_t)$
+\IND$X' := \text{: empty }m\text{-columns matrix}$
 \IND$\textbf{FOR } x' \text{ in } \Omega_t(x_t)\text{:}$
@@
 In that process, tuning correctly the $\gamma$ parameter is important.
-In the following figure, we represent two sets of actions that to respectively a gain, and a loss.
+In the following figure, we represent a set of actions that to a gain (signal = 1 here).
+The discount function that we apply is $\gamma^{\Delta-\delta}s(x^*)$.
 
-for gain and loss, display same gradient: from red to green
-SHOW IMAGE TREE WITH DIFFERENT DISCOUNT FACTORS
+\begin{figure}
+\begin{center}
+\includegraphics[width=200pt]{/Users/Thomas/git/thesis/discounts.png}
+\caption{Different discount rates affecting intermediary states}
+\end{center}
+\end{figure}
 
-We see that with a higher discount factor, a stronger emphasis is put on the final result of the set of actions.
+In the first case $\gamma=0.7$, in the second case $\gamma=0.3$.
+We see that with a larger discount factor, more emphasis is put on the final result of the set of actions.
 In some cases, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.
@@
 
-#4 Asymptotic behaviour of the algorithm
+#4 Behaviour of the algorithm
 
@@
 
-##4.1 Convergence for equivalent underlying states
+##4.1 Convergent estimates for equivalent underlying states
 

### 2016-06-06 DOESimgEQval.py
suds

@@ -2,10 +2,12 @@ import numpy as np
 import random
 import scipy
 from scipy import misc
+import scipy.misc
+from PIL import Image, ImageEnhance
 
 D = np.genfromtxt('/Users/Thomas/Dropbox/DS/thesis/data.txt', skip_footer=1)
-for i in range(0,200):
+for i in range(0,100):
 	mat = np.reshape(D[i,0:81], (-1, 9))
 	val = D[i,81]
 	name = '/Users/Thomas/git/thesis/DOESimgEQval/'+str(i)+'_'+str(val)+'.png'
-	scipy.misc.imsave(name, mat)
+	scipy.misc.toimage(mat, cmin=0, cmax=255).save(name)

### 2016-06-06 imgresize.py

@@ -1,6 +1,6 @@
 from PIL import Image, ImageEnhance
 import os
-os.chdir('/Users/Thomas/git/thesis/images6')
+os.chdir('/Users/Thomas/git/thesis/images7')
 dim = 3
 
 def normalize(arr):
@@ -15,18 +15,20 @@ def normalize(arr):
 
 for i in os.listdir(os.getcwd()):
 	if i.endswith('png'):
-		if i.startswith('X') or i.startswith('O') or i.startswith('N'):
+		if i.startswith('X') or i.startswith('O'):
+			img = Image.open(i, 'r')
+			red, green, blue, alpha = img.split()
+			img = green.resize((dim,dim), Image.ANTIALIAS).convert('LA')
+			bright = ImageEnhance.Brightness(img)
+			img = bright.enhance(1.4)
+		elif i.startswith('N'):
 			img = Image.open(i).resize((dim,dim), Image.ANTIALIAS).convert('LA')
 			bright = ImageEnhance.Brightness(img)
-			img = bright.enhance(1.3)
-		#elif i.startswith('N'):
-		#	img = Image.open(i).resize((dim,dim), Image.ANTIALIAS).convert('LA')
-		#	bright = ImageEnhance.Brightness(img)
-		#	img = bright.enhance(1.55)
-		elif i.startswith('g'):
-			img = Image.open(i).resize((dim*3,dim*3), Image.ANTIALIAS).convert('LA')
+			img = bright.enhance(1.7)
+		#elif i.startswith('g'):
+		#	img = Image.open(i).resize((dim*3,dim*3), Image.ANTIALIAS).convert('LA')
 		contrast = ImageEnhance.Contrast(img)
-		img = contrast.enhance(3)
+		img = contrast.enhance(4.5)
 
 		img.save('smalls/'+i)
 

### 2016-06-06 neural.py

@@ -26,13 +26,6 @@ def neural_train(X, y):
 	    w0 += X.T.dot(delta1)
 	    w1 += l1.T.dot(delta2)
 
-	print 'l1',l1.shape
-	print 'l2',l2.shape
-	print 'delta1',delta1.shape
-	print 'delta2',delta2.shape
-	print 'w1',l1.shape
-	print 'w0',w0.shape
-
 	return (w0, w1, l2)
 
 def neural_test(X, w0, w1):

### 2016-06-06 tic19.py

@@ -24,7 +24,7 @@ from PIL import Image, ImageEnhance, ImageDraw, ImageOps
 from lasagne import layers as lasagne, nonlinearities as nl
 import glob
 
-smalls = '/Users/Thomas/git/thesis/images6/smalls'
+smalls = '/Users/Thomas/git/thesis/images7/smalls'
 os.chdir(smalls)
 Xs = []
 Os = []
@@ -81,19 +81,18 @@ class Learner:
         s.player = player
         s.gamehist = []
         s.traced = False
+        s.net = None
         
         if player == 1:
             if alpha == None: s.alpha = 0.8
             else: s.alpha = alpha
-            if epsilon == None: s.epsilon = 0.99
+            if epsilon == None: s.epsilon = 0.99 #sqrt(other epsilon)
             else: s.epsilon = epsilon
         else: 
             if alpha == None: s.alpha = 0.8
             else: s.alpha = alpha
-            if epsilon == None: s.epsilon = 0.01
+            if epsilon == None: s.epsilon = 0.0 #propto size data.txt (http://stackoverflow.com/questions/2104080/how-to-check-file-size-in-python)
             else: s.epsilon = epsilon
-
-        s.net = None
 
     def nn_train(s, X, y):
         n = X.shape[0]
@@ -161,7 +160,13 @@ class Learner:
         row1 = np.hstack((s.stateobs(state)[0][0], s.stateobs(state)[1][0], s.stateobs(state)[2][0]))
         row2 = np.hstack((s.stateobs(state)[3][0], s.stateobs(state)[4][0], s.stateobs(state)[5][0]))
         row3 = np.hstack((s.stateobs(state)[6][0], s.stateobs(state)[7][0], s.stateobs(state)[8][0]))
-        return np.vstack((row1, row2, row3))
+        rows = np.vstack((row1, row2, row3))
+        #change the brightness randomly:
+        rows += random.sample(range(-100,0), 1)
+        #change the contrast randomly:
+        rows += random.sample(range(-100,0), 1)
+        rows[rows > 255] = 0 ; rows[rows < 0] = 0
+        return rows
 
     def value(s, state, action): #measures the gain after a particular step
         "Assumption: Game has not been won by other player"
@@ -292,9 +297,9 @@ class Game:
             img = ImageOps.autocontrast(img,ignore=range(0,135)+range(230,256))
             img = ImageOps.mirror(img)
             bright = ImageEnhance.Brightness(img)
-            img = bright.enhance(1.5)
[…]

### 2016-06-10 paper.Rmd
suds
YEYEYEYE

@@
 
+We can observe a similarity between the backpropagation mechanism of artificial neural networks. There is an activation and a backprogpagation of the values.
+
 The use of this simple sensor is made possible by the use of discounted
@@
 
+The chosen regressor is an artificial neural network.
+I chose this type of regressor as I believe it supports the complex relationship between the pixels representing a state and the value of this state.
+A sufficiently complex structure is thus able to handle the very large number of ways to represent a game combined with the many game combinations.
+In this implementation, the network is has the following form:
+PICTURE NN
+
 The stochastic state $x_t$ is resulting from another player's move represented on an image as seen above.
@@
 
-The chosen regressor is an artificial neural network.
-I chose this type of regressor as I believe it supports the complex relationship between the pixels representing a state and the value of this state.
-A sufficiently complex structure is thus able to handle the very large number of ways to represent a game combined with the many game combinations.
-In this implementation, the network is has the following form:
-PICTURE NN
+I implemented an adaptative randomness factor. It has the very *ad hoc* form 
+$$\epsilon = 50000/log_{10}(S)^6,$$
+where $S$ is the size of the data in bytes.
+I chose this formula so it equals 99% for a data corresponding to 10000 games.
+The other learner is penalized and has $\epsilon = 100000/log_{10}(S)^6$ so our learner has time to learn how to win.
 
@@
 
+###5.4 Improvements
+
+Convolutional layer
+
+Detect all illegal moves
+
 #6 Conclusion

### 2016-06-10 tic20.py

@@ -1,0 +1,548 @@
+#TO DO:
+    #IMG DISTORT: http://stackoverflow.com/questions/14177744/how-does-perspective-transformation-work-in-pil
+    #convolutional
+    #best: danielnouri.org/notes/2014/12/17/using-convolutional-neural-nets-to-detect-facial-keypoints-tutorial/
+    #One advantage is that training converges much faster; maybe four times faster in this case. The second advantage is that it also helps get better generalization; pre-training acts as a regularizer. 
+
+from __future__ import division
+from random import sample
+import datetime
+import random
+import csv
+import numpy as np
+from numpy import diag
+import cPickle
+import pickle
+import time
+import contextlib
+import os
+import sys
+import scipy
+from scipy import misc
+from sknn.mlp import Regressor, Layer, Convolution
+from PIL import Image, ImageEnhance, ImageDraw, ImageOps
+from lasagne import layers as lasagne, nonlinearities as nl
+import glob
+
+smalls = '/Users/Thomas/git/thesis/images7/smalls'
+os.chdir(smalls)
+Xs = []
+Os = []
+Ns = []
+maxi = 0
+pred = 0
+who = 'none'
+dim = 3
+selfplays = 1000
+for i in os.listdir(smalls):
+    if i.startswith('X'):
+        Xs.append(scipy.misc.imread(i, flatten=True))
+    elif i.startswith('O'):
+        Os.append(scipy.misc.imread(i, flatten=True))
+    elif i.startswith('N'):
+        Ns.append(scipy.misc.imread(i, flatten=True))
+os.chdir('/Users/Thomas/git/thesis')
+
+class State(np.ndarray):
+    symbols = {0: "_", 1: "X", 2: "O"}
+    
+    #3x3 array of zeros
+    def __new__(subtype): 
+        arr = np.zeros((3,3), dtype=np.int8)
+        return arr.view(subtype)
+
+    def __hash__(s): 
+        flat = s.ravel() #array as vector
+        code = 0
+        for i in xrange(9): code += pow(3,i) * flat[i] #pow(3,i) = 3^i
+        return code
+
+    def won(s, player): #all possibilities of winning
+        x = s == player
+        return np.hstack( (x.all(0), x.all(1), diag(x).all(), diag(x[:,::-1]).all()) ).any()
+
+    def full(s):
+        return (s != 0).all()
+
+    def __str__(s): #fill the board with current state
+        out = [""]
+        for i in xrange(3):
+            for j in xrange(3):
[…]

### 2016-06-13 paper2.tex
tic20

--- 2016-05-31 paper.Rmd
@@
----
-title: "Automated optimization of a course of actions in a stochastic environment"
-author: "Thomas Vicente"
-output: pdf_document
-fontsize: 12pt
-header-includes: \usepackage{bm}
+\documentclass[letterpaper, 10.5 pt, conference]{ieeeconf}
+\IEEEoverridecommandlockouts
+\overrideIEEEmargins
+
+\usepackage{bm}
 \usepackage{bbm}
-         \usepackage{algorithmic}
+\usepackage{algorithm,algpseudocode}
 \usepackage{algorithm}
+\usepackage{graphicx}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
 
-abstract: "I propose a principle allowing an artificial learning entity to find a right course of action possessing (like we all do) only an imperfect set of information on the current state of the world.
-I try to desmontrate the practicality and the generality of such a principle."
+\newcommand{\B}[1]{\mathbf{#1}}
 
-keywords: {reinforcement learning, neural networks, artificial intelligence}
----
+\title{\LARGE \bf
+Automated consistent course of actions in a stochastic environment
+}
 
-#1 Introduction
-A complete exercise would be to build an autonomous entity that is able to take long term oriented decisions based on information coming from the phisical reality.
+\author{Thomas Vicente$^{1}$% <-this % stops a space
+\thanks{*Candidate for the Master in Data Science at the Barcelona Graduate School of Economics.}% <-this % stops a space
+}
 
-This entity needs two abilities: learning from its own actions and recognising the current state's patterns.
-The first ability is typically made possible by a Reinforcement Learning (RL) mechanism.
-The second ability is done by regressors or classifiers powerful enough to adapt to the entity's sensorial features.
-In particular, if the entity is equipped by a visual sensor, neural networks might be a good idea.
+\begin{document}
 
-##1.1 Learning from its own actions
+\maketitle
+\thispagestyle{empty}
+\pagestyle{empty}
 
-RL emerges easily from other Machine Learning techniques as the training is done by acting.
-It uses training information that *evaluates* the actions taken rather than *instructs* by giving correct actions are faced repeatedly with a choice among n different options, or actions.
-Such setting has its roots in the *n-armed bandit problem* where after each choice you receive a numerical reward chosen from a stationary probability distribution that depends on the selected action.
+\begin{abstract}
+
+I propose an algorithm allowing the automated learning of the best course of actions in a stochastic environment.
+It notably includes a signal function that should allow more flexible implementations than current approximation-based reinforcement learning algorithms.
+This function triggers the computation of discounted action values for each learning episode.
+The paper contains an application, a tic-tac-toe robot player recognizing underlying games from photographs via a neural network.
+
+\end{abstract}
+
+\section*{Acknowledgment}
+
+I would like to thank my advisor, Professor Gábor Lugosi, for his help and for making himself available.
+I would also like to thank Gergely Neu, post-doctoral fellow at the Artificial Intelligence Research Group from the Pompeu Fabra University, for sharing his knowledge on reinforcement learning.
+
+\section{Introduction}
+
+A complete exercise would be to build an autonomous entity that is progressively able to take long term-oriented decisions based on information coming from real world sensors.
+This entity needs two abilities: learning from its own actions and recognizing useful patterns in the current state.
+The first ability typically corresponds to algorithms from the reinforcement learning (RL) field.
+The second ability requires regressor or classifier function that are able to handle the complexity of stochastic state's patterns.
+
+\subsection{Learning by action}
+
+RL emerges easily from other Machine Learning techniques as the training data is generated thanks to the actions of the learning entity.
+Thus, the \textit{learner} uses training information that evaluates the actions taken rather than instructs by giving correct actions are faced repeatedly with a choice among $n$ different options, or actions.
+Such setting has its roots in the \textit{n-armed bandit problem} where after each choice you receive a numerical reward chosen from a stationary probability distribution that depends on the selected action.
 The objective is to maximize the expected total reward over all the episodes.
 The learner evolves in a particular state and records what actions lead to a maximal gain.
-In most cases the goal is to maximise a long term gain by taing a correct course of actions.
+In most cases the goal is to maximize a long term gain by taking a correct course of actions.
 These policies can be more or less explicit.
-I our case, the proposed algorithm would be qualified as *off-policy*.
+I our case, the proposed algorithm would be qualified as \textit{off-policy}.
 
-An RL learner is thus particularly dependent on a state signal that succeeds in retaining all relevant information, a *Markov* state.
+An RL learner is thus particularly dependent on a state signal that succeeds in retaining all relevant information, a \textit{Markov} state.
 For example, a checkers position (the current configuration of all the pieces on the board) would serve as a Markov state because it summarizes everything important about the complete sequence of positions that led to it.
@@
 A reinforcement learning task that satisfies the Markov property is called a Markov Decision Process (MDP).
-If the state and action spaces are finite, it is possible the MDP framework notably formalizes the probability that a state occurs given a previous state an action, and the probability of reward after the new state occured.
+If the state and action spaces are finite, it is possible the MDP framework notably formalizes the probability that a state occurs given a previous state an action, and the probability of reward after the new state occurred.
 
-##1.2 Pattern recognition
+\subsection{Pattern recognition}
 
-Equipping a reinforcement learner with an approximator function has two benefits: the ability to provide an evaluation of an infinite set of states and recognise the a finite set of states having an infinite set of possible representations.
-In the example below, as humans we are able to recognise the same underlying situation and take action: if I learner how to make spaghettis at home, I am able to reproduce it at my friend's place.
-In a similar manner, we would like our learner to overcome its imperfect sensors to learn how to behave in the long term to maximize its gain.
+Equipping the learner with an approximation function enables it to evaluate an infinite set of states but also a finite set of underlying states having an infinite set of representations.
+In Figure~\ref{fig:surf}, I illustrate the fact that humans are able to recognize the same underlying situation and take the correct course of actions.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.35]{"surf"}
+\caption{Waves are a stochastic process but, with training, we are able to act consistently for a new wave}
+\label{fig:surf}
+\end{center}
+\end{figure}
+
+In a similar manner, we would like our learner to overcome the imperfect information it receives and learn how to behave in the long term to maximize its gain.
 This is made possible notably by regressor and classifier functions that take particularly well advantage of naturally generated data such as images and sounds.
 
-A modified version of the classical tic-tac-toerobot will provide a good example for the principle we seek to study.
+A modified version of the classical tic-tac-toe robot will provide a good example for the principle we seek to study.
 
-TWO IMAGES, caption: after enough games, the learner should be able to play the same way after vizualising either of these images
-
-#2 Litterature review
+\section{Literature review}
 
@@
 
-A this is where it differs for example from the State-Action-Reward-State-Action aglorithm (SARSA).
+A this is where it differs for example from the State-Action-Reward-State-Action algorithm (SARSA).
 It uses approximations to update the value of a whole set of actions, a policy.
@@
 It notably update the parameters of the approximation function after each movement thanks to a change in value from $t-1$ to $t$, known as a temporal difference.
-The update incldues a gradient term, compute from the difference between the state's value approximation and the real state's value.
+The update includes a gradient term, compute from the difference between the state's value approximation and the real state's value.
 
@@
 And as the learner makes better moves, the neural network is fed with more useful information.
-And thanks to the so called "Experience replay", when training the network, random minibatches from the replay memory are used instead of the most recent transition.
+And thanks to the so called "Experience replay", when training the network, random mini-batches from the replay memory are used instead of the most recent transition.
 Such efficient implementation makes the DQN very adaptive to different tasks, as demonstrated for the learning of multiple arcade games while using the same network architecture.
@@
 That would be thanks to removing the assumption that the state's values are permanently monitored during the training phase.
-This is possible by the implementation of a *signal function*.
+This is possible by the implementation of a \textit{signal function}.
 Its role is to detect a preset objective signal thanks to a deterministic function or a well tuned sensor.
-The aim is notably to allow the algorithm to be practial when evaluating non-idealized states - unlike video game screenshots - involving stochastic perturbations - such as real world pictures or videos.
+The aim is notably to allow the algorithm to be practical when evaluating non-idealized states - unlike video game screen shots - involving stochastic perturbations - such as real world pictures or videos.
 
-#3 Proposed algorithm
+\section{Algorithm}
 
-##3.1 Description
-To describe the proposed principle, we are going to replicate chronologically an episode of the algorithm in action.
+\subsection{Description}
 
-We call the learning algorithm the *learner*. The learner's aim is to automate a course of actions $x^*$ observing at $t$ a stochastic state $x_t$. Notice that $x^*$ is also equivalent to a new state
+I am going to describe chronologically what happens during an episode of the algorithm in action.
+Note that all the objects containing $X$ or $x$ in their names are $m$-dimensional objects, where $m$ is the number of features representing the considered states.
 
-At each episode, the learner is facing different choices of actions.
-The space of possible actions is called $\Omega$.
-It can be a discrete or continuous space.
-It can be set deterministically ; for instance, in the context of a game, there is a well determined space of actions that a human can tune in advance.
-It also can be infinite ; for instance, a mobile robot can explore the real 3D world in almost infinite ways.
+The learner observes a new stochastic state $\B{x_t(p_t)}$ - note that this stochastic state can be affected by previous moves from our learner.
+It then simulates actions within the finite set of possible actions $\Omega(\B{x_t})$.
+$\Omega(\B{x_t})$ can be set deterministically.
+For instance, in the context of a game, there is a well determined set of actions that a human can tune in advance.
+It also can be a subset of an infinite set of actions.
+For instance, a mobile robot can explore the real 3D world in almost infinite ways.
+In that case, the set of possible actions has to be restricted somehow, possibly by the time allowed to go in particular direction.
 
-The choice, with probability $1-\epsilon$, which corresponds to a non-exploratory move, is called $x^*$, the maximum over a set of predictions.
-These predictions correspond to a regressor function that we describe below.
-These predictions are made over $X'$, a set of simulated actions belonging to $\Omega$.
-$x^*$ is then added to a training set, $X$.
+All of these simulated actions are stored in a temporary matrix $\B{X'}$.
+Each observation in this matrix is going to be evaluated by the a approximation function $f(\B{w^*},\B{X'})$.
+This function can be a regressor or a classifier which weights $w^*$ are randomly initialized before the first episode.
+It is going to return approximations of the current state's value.
+This value is proportional to the learner's closeness to its objective.
+Among the simulated states, we will then return with probability $1-\epsilon$ the state which has the highest approximated value, $\B{x^*}$.
+With probability $\epsilon$, we will randomly pick a state from $\B{X'}$.
+If both cases, we append the returned state to the training set $\B{X}$.
 
-Now, we are assuming that the learner posesses at least one sensor allowing it to get an objective signal.
-This signal, $s(x^*)$, is a function of the state generated by the learner's last move.
-It is conditional on the current state and takes positive value if a *gain* is sensed, and negative value if a *loss* is sensed.
+The learner possesses a function $s(\B{x^*})$.
+It returns different values when a particular signal in the state generated by the learner's last move, $\B{x^*}$, is sensed.
+The value is positive if a \textit{gain} is sensed, and negative value if a \textit{loss} is sensed.
 At least one value should be potentially triggered.
 A typical example for $[gain;loss]$ would be respectively $[1;-1]$.
-They should be set so they are proportional to the learner's closeness to its objective.
+While this signal is off, the learner counts the number of actions since the signal was last triggered.
+When the signal is triggered, the current episode is over.
+Each action of the episode receives a value given by the discount function $d(\gamma, \Delta, \delta, s(x^*))$ where $\gamma$ is a discount rate, $\Delta$ is the number of actions in the episode and $\delta$ is the index of the action in the episode.
+For instance, the last action in the episode has $\delta=\Delta$.
+The function should insure that actions that are closer in time from the triggered signal get values that are closer from $s(\B{x^*})$.
+Each value of the episode are then appended to the training target vector $\B{y}$.
 
-While this signal is off, the learner counts the number of actions since the signal was last triggered.
-If this signal is triggered, the current episode is over and each action since the last signal was triggered is receiving a value given by the function $d(\gamma, \Delta, \delta, s(x^*))$.
-$\gamma$ is a discount rate, $\Delta$ the number of actions in the episode and $\delta$ the index of an action in the episode (the first action in the episode has $\delta=1$).
-So the closer an action was from the triggered signal, the larger the magnitude of its value will be.
-Each value corresponding to all the last actions are appended to the training target vector $y$.
-I realize this is only a particular way to discount these values, and different ways might be considered for different applications.
+We then get a training sample on which are going to apply our approximation function $f(\B{w}, \B{X})$ which this time minimizes the loss $l(f(\B{w}, \B{X}),\B{y})$.
+From this process, we are going to retain set of parameters corresponding to the minimal loss to update $w^*$.
+The updated $w^*$ is going to be used in the next episode in order to predict the value of simulated actions.
 
-We then get a first training sample on which are going to apply a regressor function $f()$.
-Its aim is to minimize the loss $l(f(w,X),y)$.
-The parameter set corresponding to the minimal loss is $w^*$.
-It is going to be used in the next episode in order to predict the value of simulated actions
+Algorithm~\ref{alg:algo} is a formal version of the proposed algorithm.
 
-Here is a formal version of the algorithm:
+\begin{algorithm}
+\caption{Approximation-based reinforcement learning with signal function}
+\label{alg:algo}
+\begin{algorithmic}[1]
 
-\newpage
+\State $m$ : number of features
+\State $\B{X}=\B{X_{\o}}$: empty $m$-columns matrix
+\State $\B{y}=\B{y_{\o}}$: empty target vector
+\State $\B{w^*}=\B{w_0}$: random weights for the regressor
+\State $\gamma \in [0,1]$  discount rate
+\State $\Delta = 0$ : number of actions
+\State $f()$ : regressor function
+\State $p_t(\B{x^*})$: stochastic process, function of the last existing chosen state
+\State $\B{X'}$: $m$-columns simulated states matrix
+\State $\B{x_t(p_t)} \in {\rm I\!R^m}$: current state vector
+\State $\Omega_t(\B{x_t})$ : set of possible actions
+\State $\epsilon$: randomness rate
+\State $s(\B{x^*}) \in {\rm I\!R}$: signal value potentially triggered by the modified state
 
-\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
-\begin{algorithm*}
-\caption{}
-\begin{algorithmic}
-\STATE$\text{Initialized variables:}$
-\IND$X\text{: empty }m\text{-columns feature matrix}$
-\IND$y\text{: empty target vector}$
-\IND$w^*=w_0\text{: random weights for the regressor}$
-\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
-\IND$\Delta = 0 \text{: number of actions}$
-\STATE$\text{Other variables:}$
-\IND$f() \text{: regressor function}$
-\IND$m \text{: number of features}$
-\IND$p_t(x^*) \text{: stochastic process, function of the last existing chosen state}$
-\IND$X'\text{: }m\text{-columns simulated states matrix}$
-\IND$x_t(p_t) \in {\rm I\!R^m} \text{: current state vector}$
-\IND$\Omega_t(x_t) \text{: set of possible actions}$
-\IND$\epsilon \text{: randomness rate}$
-\IND$s(x^*) \in {\rm I\!R} \text{: signal value potentially triggered by the modified state}$
-\STATE$\textbf{WHILE } \text{the learning process is on:}$
-\IND$\text{record }x_t(p_t)$
-\IND$X' := \text{: empty }m\text{-columns matrix}$
-\IND$\textbf{FOR } x' \text{ in } \Omega_t(x_t)\text{:}$
-\IND[2]$\text{append } x' \text{ to } X'$
-\IND$\text{With probability } 1-\epsilon$
-\IND[2]$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
-\IND$\text{With probability } \epsilon$
-\IND[2]$x^*:=\text{random sample }x' \in X'$
-\IND$\text{append } x^* \text{ to } X$
-\IND$\textbf{IF } s(x^*) \text{ != } 0 \text{ :}$
-\IND[2]$\textbf{FOR } \delta \text{ in } 1\text{:}\Delta \text{:}$
-\IND[3]$value = d(\gamma, \Delta, \delta, s(x^*))$
-\IND[3]$\text{append } value \text{ to } y$
-\IND[2]$w^* := \arg\min_{w}(l(f(w,X),y))$
-\IND[2]$\Delta := 0$
-\IND$\textbf{ELSE} \text{:}$
-\IND[2]$\Delta \text{ += } 1$
+\While{the learning process is on}
+  \State record $\B{x_t(p_t)}$
+    \State $\B{X'} :=\B{X'_{\o}}$ empty $m$-columns matrix
+    \For{$\B{x'}$  in  $\Omega(\B{x_t})$}
+      \State append  $\B{x'}$  to  $\B{X'}$
+    \EndFor
+    \State with probability  $1-\epsilon$
+    \State \hspace{0.5cm} $\B{x^*}:=\arg\max_{\B{x'} \in \B{X'}} \in f(\B{w^*},\B{X'})$
+    \State with probability  $\epsilon$
+    \State \hspace{0.5cm} $\B{x^*}$ := random sample $\B{x'} \in \B{X'}$
+    \State append $\B{x^*}$ to $\B{X}$
+    \If{$s(\B{x^*})$  !=  $0$}
+      \For{$\delta$  in  $1:\Delta$}
+        \State $v$ = $d(\gamma, \Delta, \delta, s(\B{x^*}))$
+        \State append  $v$  to  $\B{y}$
+       \EndFor
+      \State $\B{w^*} := \arg\min_{\B{w}}(l(f(\B{w},\B{X}),\B{y}))$
+      \State $\Delta := 0$
+    \Else
+      \State $\Delta$ += $1$
+    \EndIf
+\EndWhile
+
 \end{algorithmic}
-\end{algorithm*}
+\end{algorithm}
 
-\newpage
+\subsection{Remarks}
 
-##3.2 Remarks
-
-###3.2.1 Regressor function
-
-Depending of the sensor the learner has at his disposal, some regressor functions can be more or less suited.
-
-If the possible states form a linear function of the target value, a linear regressor function can be applied.
-But the states can be strongly affected by stochastic processes.
-For structured data, non-parametric tree-based methods might be considers.
-Our typical case is when this happens with unstructured data, such as images, sounds and texts.
-Then, neural networks might be considered.
-
-This last case is typically mimicking how human would act if faced with a particular situation.
-The challenge is here to recognise equivalent underlying states thanks to pattern recognition.
-And considering the algorithm described above, $w$, would be the set of a neural network's hidden layers.
+\subsubsection{Regressor function}
 
 One interesting aspect is the role of the regressor function.
-Contrarily to typical applications, the functions's task is utimately not to predict an action's value with maximum accuracy, even though it is often a dependency, but rather to give the "real value" of an action.
+Contrarily to typical applications, the functions's task is utimately not to predict an action's value with maximum accuracy, even though it is often related to what we want, but rather to give the "real value" to an action.
 Imagine for instance a game where two intermediary states might lead to both winning and losing.
-Such states will get a non-tendencial value from the regressor function, which is a good thing.
+Such states will get a non-tendential value from the function, which is a good thing.
 
-###3.2.2 Signal function and discounted values
+Even if the approximation function's task is this time to unveil the real value of each state, the choice for the right function is still clearly very important.
+It should notably depend on the complexity of the patterns that we wish the learner to discover, which is often proportional to the complexity of the considered $m$-dimensional states.
 
-The signal function allows for minimal assumptions on the basic abilities of the learner.
-While most approximation-based reinforcement learning algorithms suppose the permanent presence of a value monitoring function, here we only assume the presence of a discrete, time-independent sensor triggering one or more values.
+Obviously, if these states approximately form a linear function of the target value, a linear approximation function can be applied.
+But if the states are strongly affected by stochastic processes, other functions could be considered.
 
-The use of this simple sensor is made possible by the use of discounted
-values.
-Even if not every action is monitored during an episode, each step still takes a value at the end of the episodes.
+Typically in this paper, we consider real-world representations as states.
+In other words, our input could be include images, sounds and texts.
+In that case, we might consider using neural networks which are often celebrated as a particularly accurate family of algorithms for such data.
+Using neural networks, we can see our learner as an imitator of how humans would act when facing the task in question.
+The challenge is then to recognize equivalent underlying states thanks to pattern recognition.
+Finally note, considering the algorithm described above, that $\B{w}$ would be the set of a neural network's hidden layers.
+
+\subsubsection{Signal function and discounted values}
+
+The signal function allows to reduce the assumptions on the basic abilities of the learner.
+While most approximation-based reinforcement learning algorithms suppose the permanent presence of a value-monitoring function, here we only assume the presence of a discrete, time-independent sensor triggering one or more values.
+
+The use of discounted values enables each step to get a value at the end of each episodes.
 In that process, tuning correctly the $\gamma$ parameter is important.
-In the following figure, we represent a set of actions that to a gain (signal = 1 here).
-The discount function that we apply is $\gamma^{\Delta-\delta}s(x^*)$.
+In Figure~\ref{fig:discounts}, we represent a set of actions that to a gain - $s(\B{x^*} = 1$ here, and the discount function that we apply is $\gamma^{\Delta-\delta}s(\B{x^*})$.
 
@@
 \begin{center}
-\includegraphics[width=200pt]{/Users/Thomas/git/thesis/discounts.png}
+\includegraphics[scale=.5]{"discounts"}
 \caption{Different discount rates affecting intermediary states}
+\label{fig:discounts}
 \end{center}
@@
 We see that with a larger discount factor, more emphasis is put on the final result of the set of actions.
-In some cases, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.
+If relevant, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.
 
 With this function, triggering a new signal can take time.
-But the learner can still look for best discounted value according to the current state.
-We then benefit from something common to RL algorithms: the use of intermediary objectives.
+But the learner still looks for best discounted value according to the current state.
+Thus, with less assumptions on the monitoring of the real value of the states, we still benefit from something common to RL algorithms: the discovery and use of intermediary objectives.
 
-#4 Behaviour of the algorithm
+\section{Behaviour of the algorithm}
 
-include idea RL+regressor (Gabor)
-
-##4.1 Convergent estimates for equivalent underlying states
+\subsection{Convergent estimates for equivalent underlying states}
 
@@
 
-##4.2 Long term gain maximization
+\subsection{Long term gain maximization}
 
@@
 
-#5 Implementaion for a "real-world" tic-tac-toe game
+\section{Implementaion for a "real-world" tic-tac-toe game}
 
@@
 Notably, when playing against a human, the information available to the robot corresponds to an image of the type:
-PICTURES OF VISION
-caption: pictures of the same underlying state
+
+TWO IMAGES
+caption: after enough games, the learner should be able to play the same way after visualizing either of these images
+
 This is thus a "stochastic representation" from a finite set of states. The stochastic processes affecting this representation are notably the varying handwrittings and the environment's brightness.
@@
 
-##5.1 Particularities of the implementation
+\subsection{Particularities of the implementation}
+
+The chosen regressor is an artificial neural network detailed in Figure~\ref{fig:neuralnet}.
+I chose this type of regressor as I believe it supports the complex relationship between the pixels representing a state and the value of this state.
+A sufficiently complex structure is thus able to handle the very large number of ways to represent a game combined with the many game combinations.
+In this implementation, the network is has the following form:
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.25]{"NNview"}
+\caption{Other parameters are: the number of iterations = 200, the learning rate = 0.0001, the learning rule = Rmsprop which divides the learning rate for a weight by a running average of the magnitudes of recent gradients for that weight, the threshold under which the validation error change is assumed to be stable = 0.0001, the mini-batch size = 200 observations, and the number of iterations after which training should return when the validation error remains (near) constant = 10.}
+\label{fig:neuralnet}
+\end{center}
+\end{figure}
 
@@
 
-The signal function is determincally set by a human.
+The signal function is set by a human.
 It triggers the values $1$ if the learner wins, $-0.1$ if there is a draw and $-1$ if it loses or if it does an illegal move.
@@
 
-The chosen regressor is an artificial neural network.
-I chose this type of regressor as I believe it supports the complex relationship between the pixels representing a state and the value of this state.
-A sufficiently complex structure is thus able to handle the very large number of ways to represent a game combined with the many game combinations.
-In this implementation, the network is has the following form:
-PICTURE NN
+I implemented an adaptive randomness factor. It has the very *ad hoc* form 
+$$\epsilon = 50000/log_{10}(S)^6,$$
+where $S$ is the size of the data in bytes.
+I chose this formula so it equals 99\% for a data corresponding to 10000 games.
+The other learner is penalized and has $\epsilon = 100000/log_{10}(S)^6$ so our learner has time to learn how to win.
 
-##5.2 Necessary adjustments
+\subsection{Necessary adjustments}
 
@@
 This is purely for computational constraints.
-However, this does not affect in fine the performance of the learner, which is able to win up to 90% of the last phases' games.
+However, this does not affect in fine the performance of the learner, which is able to win up to 90\% of the last phases' games.
 
-##5.3 Performance analysis
+\subsection{Performance analysis}
 
@@
 
-```{r,echo=F}
-#show in grid, different epsilons for 1000 games, 10000 also?
-setwd('/Users/Thomas/git/thesis/reports/')
+INCLUDE R PLOTS
 
-plotsmthg <- function(what, nbg, random, colour, yrange){
-  games <- seq(nbg/10,nbg,nbg/10)
-  d <- read.csv(paste0(nbg,'games_epsilon',random,'.txt'), header=F)
-  plot(games, d[,what], type='l', main = paste0('Randomness of the oponent=',random), xlab='Number of games', ylab='', bty='n', col=colour,ylim=yrange)
-}
+We see that a more exploratory learner improves its winning rate faster but is then relatively limited towards the last phases.
 
-plotsmthg(1,100,0.99,'green',range(0,1))
-par(new=T)
-plotsmthg(2,100,0.99,'red',range(0,1))
-
-#plotsmthg(3,100,0.99,'black',range(3,4))
-```
-
-We see that a more eploratory learner improves its winning rate faster but is then relatively limited towards the last phases.
-
-A more exploratory oponent helps the learner to win and thus discover what are the winning strategies.
-However, making the oponent too exploratory would limit the performance of the learner when playing against a trained human.
+A more exploratory opponent helps the learner to win and thus discover what are the winning strategies.
+However, making the opponent too exploratory would limit the performance of the learner when playing against a trained human.
 
@@
 
-Notice that a learner trained against a random oponent is able, after 100 games, to play easy moves, typically the last move to win, against a human player.
+Notice that a learner trained against a random opponent is able, after 100 games, to play easy moves, typically the last move to win, against a human player.
 
@@
 
-#6 Conclusion
+\subsection{Improvements}
+
+Convolutional layer
+
+Detect all illegal moves
+
+\section{Conclusion}
 
@@
 It could also be useful in online marketing to automatically identify set of actions leading to better sales, such as the sequence of clicks of customers.
-We also can imagine learning (thanks to vision) the way natural organisms evolve in their environment to build very adaptative machines.
+We also can imagine learning (thanks to vision) the way natural organisms evolve in their environment to build very adaptive machines.
 
-#8 References
-Sutton and Barto, *Reinforcement learning: An introduction*, Vol.
-2.
-No.
-1, MIT press, 2012, available at [people.inf.elte.hu/lorincz/Files/RL_2006/SuttonBook.pdf](people.inf.elte.hu/lorincz/Files/RL_2006/SuttonBook.pdf)
+\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
+                                  % on the last page of the document manually. It shortens
+                                  % the textheight of the last page by a suitable amount.
+                                  % This command does not take effect until the next page
+                                  % so it should come on the page before the last. Make
+                                  % sure that you do not shorten the textheight too much.
 
-http://www0.cs.ucl.ac.uk/staff/D.Silver/web/Teaching_files/FA.pdf
+\begin{thebibliography}{99}
 
-https://www.cs.toronto.edu/~vmnih/docs/dqn.pdf
+\bibitem{c1} Sutton and Barto, Reinforcement learning: An introduction, Vol. 2. No. 1, MIT press, 2012
 
-http://gaips.inesc-id.pt/~fmelo/pub/melo08icml.pdf
+\bibitem{c2} $http://www0.cs.ucl.ac.uk/staff/D.Silver/web/Teaching_files/FA.pdf$
 
-*Human-level control through deep reinforcement learning* http://www.readcube.com/articles/10.1038/nature14236?shared_access_token=Lo_2hFdW4MuqEcF3CVBZm9RgN0jAjWel9jnR3ZoTv0P5kedCCNjz3FJ2FhQCgXkApOr3ZSsJAldp-tw3IWgTseRnLpAc9xQq-vTA2Z5Ji9lg16_WvCy4SaOgpK5XXA6ecqo8d8J7l4EJsdjwai53GqKt-7JuioG0r3iV67MQIro74l6IxvmcVNKBgOwiMGi8U0izJStLpmQp6Vmi_8Lw_A%3D%3D
+\bibitem{c3} https://www.cs.toronto.edu/~vmnih/docs/dqn.pdf
+
+\bibitem{c4} http://gaips.inesc-id.pt/~fmelo/pub/melo08icml.pdf
+
+\bibitem{c5} Human-level control through deep reinforcement learning
+
+\end{thebibliography}
+
+\end{document}

### 2016-06-13 tic20.py

@@ -110,14 +110,14 @@ class Learner:
                     Layer("Sigmoid", units=200),
                     Layer("Sigmoid")],
                 n_iter=200,
-                learning_rate=0.00005,
+                learning_rate=0.0001,
                 parameters=params,
                 valid_set=cv,
                 learning_rule='rmsprop',
                 f_stable=0.0001,
                 verbose=True,
                 batch_size=200,
-                n_stable=10)
+                n_stable=7)
 
         return s.net.fit(X, y)
             
@@ -221,8 +221,8 @@ class Learner:
             log_size = np.log10(data_size)
         else: log_size = 1
 
-        if s.player == 1: s.epsilon = min([1, 100000/(log_size**6)]) #we penalize the Q-learner. It would be too strong at the begining!
-        else: s.epsilon = min([1, 50000/(log_size**6)])
+        if s.player == 1: s.epsilon = min([1, 120000/(log_size**6)]) #we penalize the Q-learner. It would be too strong at the begining!
+        else: s.epsilon = min([1, 30000/(log_size**6)])
 
         split = np.random.choice(2, 1, p=[1 - s.epsilon, s.epsilon]).tolist()[0]
         if split == 1: return rc
@@ -417,14 +417,10 @@ class Selfplay:
                     init = s.state[b]
                     s.state[b] = 2
                     ratio = round((s.games+1)/selfplays, 0)
-                    if ratio <= .2:
-                        j = 4
-                    elif ratio <= .4:
-                        j = 3
-                    elif ratio <= .5:
-                        j = 2
-                    else:
-                        j = 1
+                    if ratio <= .2: j = 4
+                    elif ratio <= .4: j = 3
+                    elif ratio <= .5: j = 2
+                    else: j = 1
                     for i in random.sample(range(0,4), j):
                         rot1 = np.rot90(s.state, i)
                         s.X1 = np.append(s.X1, [s.learner.fullimg(rot1).flatten()], axis=0)
@@ -458,23 +454,17 @@ class Selfplay:
             if s.state.full() or s.state.won(1) or s.state.won(2):
                 for m in range(1, s.moves+1):
                     if s.state.won(1):
-                        if m == s.moves:
-                            s.val = 0
-                        else:
-                            s.val = (1 - s.gamma) ** m
-                    elif s.state.won(2):
-                        s.val = s.gamma ** (s.moves - m)
-                    else:
[…]

### 2016-06-24 DOESimgEQval.py
sds

@@ -4,7 +4,7 @@ import scipy
 from scipy import misc
 import scipy.misc
 from PIL import Image, ImageEnhance
-dim = 5
+dim = 3
 
 D = np.genfromtxt('/Users/Thomas/Dropbox/DS/thesis/data.txt', skip_footer=1)
 for i in range(0,100):

### 2016-06-24 imgresize.py

@@ -1,7 +1,7 @@
 from PIL import Image, ImageEnhance
 import os
 os.chdir('/Users/Thomas/git/thesis/images7')
-dim = 5
+dim = 3
 
 def normalize(arr):
     arr = arr.astype('float')
@@ -20,15 +20,15 @@ for i in os.listdir(os.getcwd()):
 			red, green, blue, alpha = img.split()
 			img = green.resize((dim,dim), Image.ANTIALIAS).convert('LA')
 			bright = ImageEnhance.Brightness(img)
-			#img = bright.enhance(1.4)
+			img = bright.enhance(1.7)
 		elif i.startswith('N'):
 			img = Image.open(i).resize((dim,dim), Image.ANTIALIAS).convert('LA')
 			bright = ImageEnhance.Brightness(img)
-			#img = bright.enhance(1.7)
+			img = bright.enhance(1.7)
 		#elif i.startswith('g'):
 		#	img = Image.open(i).resize((dim*3,dim*3), Image.ANTIALIAS).convert('LA')
 		contrast = ImageEnhance.Contrast(img)
-		#img = contrast.enhance(4.5)
+		img = contrast.enhance(6)
 
 		img.save('smalls/'+i)
 

### 2016-06-24 tic20.py

@@ -1,13 +1,12 @@
-#TO DO:
-    #IMG DISTORT: http://stackoverflow.com/questions/14177744/how-does-perspective-transformation-work-in-pil
-    #convolutional
-    #best: danielnouri.org/notes/2014/12/17/using-convolutional-neural-nets-to-detect-facial-keypoints-tutorial/
-    #One advantage is that training converges much faster; maybe four times faster in this case. The second advantage is that it also helps get better generalization; pre-training acts as a regularizer. 
+#IMG DISTORT: http://stackoverflow.com/questions/14177744/how-does-perspective-transformation-work-in-pil
+#convolutional
+#best: danielnouri.org/notes/2014/12/17/using-convolutional-neural-nets-to-detect-facial-keypoints-tutorial/
+#One advantage is that training converges much faster; maybe four times faster in this case. The second advantage is that it also helps get better generalization; pre-training acts as a regularizer. 
 
 from __future__ import division
 from random import sample
 import datetime
-import random
+import random&
 import csv
 import numpy as np
 from numpy import diag
@@ -32,7 +31,7 @@ Ns = []
 maxi = 0
 pred = 0
 who = 'none'
-dim = 5
+dim = 3
 selfplays = 1000
 for i in os.listdir(smalls):
     if i.startswith('X'):
@@ -130,7 +129,7 @@ class Learner:
         for i in xrange(3):
             for j in xrange(3):
                 #if a given position in the given state
-                #is still empty then add it as a possible action 
+                #is still empty then add it as a possible action
                 if state[i,j] == 0:
                     res.append((i,j))
         return res
@@ -158,9 +157,9 @@ class Learner:
         row3 = np.hstack((s.stateobs(state)[6][0], s.stateobs(state)[7][0], s.stateobs(state)[8][0]))
         rows = np.vstack((row1, row2, row3))
         #change the brightness randomly:
-        rows += random.sample(range(-70,0), 1)
+        #rows += random.sample(range(-40,0), 1)
         #change the contrast randomly:
-        rows += random.sample(range(-70,0), 1)
+        #rows += random.sample(range(-70,0), 1)
         rows[rows > 255] = 255 ; rows[rows < 0] = 0
         return rows
 
@@ -221,7 +220,7 @@ class Learner:
             log_size = np.log10(data_size)
         else: log_size = 1
 
[…]

### 2016-06-24 tic21.py

@@ -1,0 +1,540 @@
+#IMG DISTORT: http://stackoverflow.com/questions/14177744/how-does-perspective-transformation-work-in-pil
+#convolutional
+#best: danielnouri.org/notes/2014/12/17/using-convolutional-neural-nets-to-detect-facial-keypoints-tutorial/
+#One advantage is that training converges much faster; maybe four times faster in this case. The second advantage is that it also helps get better generalization; pre-training acts as a regularizer. 
+
+from __future__ import division
+from random import sample
+import datetime
+import random
+import csv
+import numpy as np
+from numpy import diag
+import cPickle as pickle
+import time
+import contextlib
+import os
+import sys
+import scipy
+from scipy import misc
+from sknn.mlp import Regressor, Layer, Convolution
+from PIL import Image, ImageEnhance, ImageDraw, ImageOps
+from lasagne import layers as lasagne, nonlinearities as nl
+import glob
+from sknn.platform import cpu64
+os.environ["THEANO_FLAGS"] = "mode=FAST_RUN,device=cpu,exception_verbosity=high"
+import theano
+
+smalls = '/Users/Thomas/git/thesis/images7/smalls'
+os.chdir(smalls)
+Xs = [] 
+Os = []
+Ns = []
+maxi = 0
+pred = 0
+who = 'none'
+dim = 3
+selfplays = 1000
+for i in os.listdir(smalls):
+    if i.startswith('X'):
+        Xs.append(scipy.misc.imread(i, flatten=True))
+    elif i.startswith('O'):
+        Os.append(scipy.misc.imread(i, flatten=True))
+    elif i.startswith('N'):
+        Ns.append(scipy.misc.imread(i, flatten=True))
+os.chdir('/Users/Thomas/git/thesis')
+
+class State(np.ndarray):
+    symbols = {0: "_", 1: "X", 2: "O"}
+    
+    #3x3 array of zeros
+    def __new__(subtype): 
+        arr = np.zeros((3,3), dtype=np.int8)
+        return arr.view(subtype)
+
+    def __hash__(s): 
+        flat = s.ravel() #array as vector
+        code = 0
+        for i in xrange(9): code += pow(3,i) * flat[i] #pow(3,i) = 3^i
+        return code
+
+    def won(s, player): #all possibilities of winning
+        x = s == player
+        return np.hstack( (x.all(0), x.all(1), diag(x).all(), diag(x[:,::-1]).all()) ).any()
+
+    def full(s):
+        return (s != 0).all()
+
[…]

### 2016-06-25 paper2.tex
suds

@@
 \usepackage{lmodern}
+\usepackage{amsmath}
 
@@
 I would like to thank my advisor, Professor Gábor Lugosi, for his help and for making himself available.
-I would also like to thank Gergely Neu, post-doctoral fellow at the Artificial Intelligence Research Group from the Pompeu Fabra University, for sharing his knowledge on reinforcement learning.
+I would also like to thank Gergely Neu, post-doctoral fellow at the Artificial Intelligence Research Group from the Pompeu Fabra University, for sharing his knowledge.
 
@@
 
-The proposed algorithm allows to optimize a course of actions in finite and infinite spaces. But it has especially been thought to act based on imperfectly sensed states as described above.
+The proposed algorithm posseses similarities with the concepts that we are presenting in this section.
+We also are going to show that, to my knowledge, it is a novel approach in approximation-based RL.
+This would notably be due to the algorithm's minimal requirements for control of the states' values.
 
-A this is where it differs for example from the State-Action-Reward-State-Action algorithm (SARSA).
+A close classical RL technique to the proposed algorithm, is the the State-Action-Reward-State-Action algorithm (SARSA).
 It uses approximations to update the value of a whole set of actions, a policy.
@@
 The update includes a gradient term, compute from the difference between the state's value approximation and the real state's value.
+Our algorithm however does not suppose the learning of a whole policy.
 
-The Deep Q networks (DQN) are close to what I propose.
-This consists in an off-policy algorithm using neural network approximations of the states' value to decide on an action.
+In that sense, the recent Deep Q networks (DQN) are closer to what is proposed.
+This off-policy algorithm uses neural network approximations of the states' value to decide on an action.
 And as the learner makes better moves, the neural network is fed with more useful information.
-And thanks to the so called "Experience replay", when training the network, random mini-batches from the replay memory are used instead of the most recent transition.
+And thanks to its \textit{experience replay}, when training the network, random mini-batches from the replay memory are used instead of the most recent transition.
 Such efficient implementation makes the DQN very adaptive to different tasks, as demonstrated for the learning of multiple arcade games while using the same network architecture.
+I also assume such memory device for the proposed algorithm.
+But I try to have a more flexible approach than this algorithm, which supposes the constant monitoring of the states' values.
 
-Hopefully the proposed algorithm offers even more flexibility than the evoked algorithms.
-That would be thanks to removing the assumption that the state's values are permanently monitored during the training phase.
-This is possible by the implementation of a \textit{signal function}.
-Its role is to detect a preset objective signal thanks to a deterministic function or a well tuned sensor.
-The aim is notably to allow the algorithm to be practical when evaluating non-idealized states - unlike video game screen shots - involving stochastic perturbations - such as real world pictures or videos.
+This is done thanks to a mechanism that is close to the True Online Temporal Difference Learning from Seijen et al..
+The authors introduced a formal way to assign values in a backward manner to each state of the episode, when the latter ends.
+This is done in a discounted way so states farther to the end of an episode get a value with a smaller magnitude.
+In the proposed algorithm, however, as mentioned earlier, we want to avoid using a permanent control of the states' values during the learning phase.
+It then differs from the mentioned paper as we cannot update our parameters via traces.
+Instead, the improving estimation of the new states' value is done exclusively via the approximation function.
+
+Hopefully the proposed algorithm offers a greater flexibility than the previous litterature.
+Removing the assumption that the state's values are permanently monitored during the training phase is made possible by the implementation of a \textit{signal function} coupled with a \textit{discount function}.
+The role of the first function is to detect an objective signal thanks to a deterministic function or a well tuned sensor.
+When triggered, it assigns a gain or loss-related value to the final state of the episode.
+The second function will allow to assign values to each state of a particular episode.
+This mechanism is supposed to allow a practical use of the algorithm when evaluating non-idealized states - not only video games screen shots - involving stochastic perturbations that are for instance recorded via a visual sensor.
 
@@
 
-I am going to describe chronologically what happens during an episode of the algorithm in action.
+We can summarise our setting as a decision process possibly lacking the Markov property. It has the form
+$$\langle \mathcal{X},\Omega, P(\B{x_c},\B{x_{t+1}}), \B{y},s(\B{x_c}),d(\gamma, \Delta, \delta, s(\B{x_c})) \rangle$$
+where $\mathcal{X}$ is a set of finite underlying states, $\Omega$ is the finite set of possible actions, $P(\B{x_c})$ is the probability that the process moves to a new state $\B{x_{t+1}}$ potentially depending on numerous previous states in the episode, $s()$ is the signal function and d() is a the discount function.
+
+I am going to describe more details and chronologically what happens during an episode of the algorithm in action.
 Note that all the objects containing $X$ or $x$ in their names are $m$-dimensional objects, where $m$ is the number of features representing the considered states.
 
-The learner observes a new stochastic state $\B{x_t(p_t)}$ - note that this stochastic state can be affected by previous moves from our learner.
+The learner observes a new stochastic state $\B{x_t}(p_t)$ - note that this stochastic state can be affected by previous moves from our learner.
 It then simulates actions within the finite set of possible actions $\Omega(\B{x_t})$.
@@
 All of these simulated actions are stored in a temporary matrix $\B{X'}$.
-Each observation in this matrix is going to be evaluated by the a approximation function $f(\B{w^*},\B{X'})$.
-This function can be a regressor or a classifier which weights $w^*$ are randomly initialized before the first episode.
+Each observation in this matrix is going to be evaluated by the a approximation function $f(\B{w},\B{X'})$.
+This function can be a regressor or a classifier which weights $w$ are randomly initialized before the first episode.
 It is going to return approximations of the current state's value.
 This value is proportional to the learner's closeness to its objective.
-Among the simulated states, we will then return with probability $1-\epsilon$ the state which has the highest approximated value, $\B{x^*}$.
-With probability $\epsilon$, we will randomly pick a state from $\B{X'}$.
+Among the simulated states, we will then return the chosen $\B{x}$, $\B{x_c}$.
+With probability $1-\epsilon$ it corresponds to the state which has the highest approximated valu and with probability $\epsilon$, it will be a randomly chosen state from $\B{X'}$.
 If both cases, we append the returned state to the training set $\B{X}$.
 
-The learner possesses a function $s(\B{x^*})$.
-It returns different values when a particular signal in the state generated by the learner's last move, $\B{x^*}$, is sensed.
+The learner possesses a function $s(\B{x_c})$.
+It returns different values when a particular signal in the state generated by the learner's last move, $\B{x_c}$, is sensed.
 The value is positive if a \textit{gain} is sensed, and negative value if a \textit{loss} is sensed.
@@
 When the signal is triggered, the current episode is over.
-Each action of the episode receives a value given by the discount function $d(\gamma, \Delta, \delta, s(x^*))$ where $\gamma$ is a discount rate, $\Delta$ is the number of actions in the episode and $\delta$ is the index of the action in the episode.
+Each action of the episode receives a value given by the discount function $d(\gamma, \Delta, \delta, s(\B{x_c}))$ where $\gamma$ is a discount rate, $\Delta$ is the number of actions in the episode and $\delta$ is the index of the action in the episode.
 For instance, the last action in the episode has $\delta=\Delta$.
-The function should insure that actions that are closer in time from the triggered signal get values that are closer from $s(\B{x^*})$.
+The function should insure that actions that are closer in time from the triggered signal get values that are closer from $s(\B{x_c})$.
 Each value of the episode are then appended to the training target vector $\B{y}$.
@@
 We then get a training sample on which are going to apply our approximation function $f(\B{w}, \B{X})$ which this time minimizes the loss $l(f(\B{w}, \B{X}),\B{y})$.
-From this process, we are going to retain set of parameters corresponding to the minimal loss to update $w^*$.
-The updated $w^*$ is going to be used in the next episode in order to predict the value of simulated actions.
+From this process, we are going to retain set of parameters corresponding to the minimal loss to update $w$.
+The updated $w$ is going to be used in the next episode in order to predict the value of simulated actions.
 
@@
 \begin{algorithm}
-\caption{Approximation-based reinforcement learning with signal function}
-\label{alg:algo}
+
 \begin{algorithmic}[1]
 
-\State $m$ : number of features
-\State $\B{X}=\B{X_{\o}}$: empty $m$-columns matrix
-\State $\B{y}=\B{y_{\o}}$: empty target vector
-\State $\B{w^*}=\B{w_0}$: random weights for the regressor
-\State $\gamma \in [0,1]$  discount rate
-\State $\Delta = 0$ : number of actions
-\State $f()$ : regressor function
-\State $p_t(\B{x^*})$: stochastic process, function of the last existing chosen state
-\State $\B{X'}$: $m$-columns simulated states matrix
-\State $\B{x_t(p_t)} \in {\rm I\!R^m}$: current state vector
-\State $\Omega_t(\B{x_t})$ : set of possible actions
-\State $\epsilon$: randomness rate
-\State $s(\B{x^*}) \in {\rm I\!R}$: signal value potentially triggered by the modified state
-
+\vspace{3mm}
 \While{the learning process is on}
-  \State record $\B{x_t(p_t)}$
+  \State record $\B{x_t}(p_t)$
     \State $\B{X'} :=\B{X'_{\o}}$ empty $m$-columns matrix
@@
     \State with probability  $1-\epsilon$
-    \State \hspace{0.5cm} $\B{x^*}:=\arg\max_{\B{x'} \in \B{X'}} \in f(\B{w^*},\B{X'})$
+    \State \hspace{0.5cm} $\B{x_c}:=\arg\max_{\B{x'} \in \B{X'}} \in f(\B{w},\B{X'})$
     \State with probability  $\epsilon$
-    \State \hspace{0.5cm} $\B{x^*}$ := random sample $\B{x'} \in \B{X'}$
-    \State append $\B{x^*}$ to $\B{X}$
-    \If{$s(\B{x^*})$  !=  $0$}
+    \State \hspace{0.5cm} $\B{x_c}$ := random sample $\B{x'} \in \B{X'}$
+    \State append $\B{x_c}$ to $\B{X}$
+    \If{$s(\B{x_c})$  !=  $0$}
       \For{$\delta$  in  $1:\Delta$}
-        \State $v$ = $d(\gamma, \Delta, \delta, s(\B{x^*}))$
+        \State $v$ = $d(\gamma, \Delta, \delta, s(\B{x_c}))$
         \State append  $v$  to  $\B{y}$
        \EndFor
-      \State $\B{w^*} := \arg\min_{\B{w}}(l(f(\B{w},\B{X}),\B{y}))$
+      \State $\B{w} := \arg\min_{\B{w}}(l(f(\B{w},\B{X}),\B{y}))$
       \State $\Delta := 0$
@@
 \end{algorithmic}
+
+\caption{}
+\label{alg:algo}
+
+\vspace{3mm}
+\begin{small} $m$: number of features, $\B{X}=\B{X_{\o}}$: empty $m$-columns matrix, $\B{y}=\B{y_{\o}}$: empty target vector, $\B{w}=\B{w_0}$: random weights for the regressor, $\gamma \in [0,1]$  discount rate, $\Delta = 0$: number of actions, $f()$: regressor function, $p_t(\B{x_c})$: stochastic process, function of the last existing chosen state, $\B{X'}$: $m$-columns simulated states matrix, $\B{x_t}(p_t) \in {\rm I\!R^m}$: current state vector, $\Omega_t(\B{x_t})$: set of possible actions, $\epsilon$: randomness rate, $s(\B{x_c}) \in {\rm I\!R}$: signal value potentially triggered by the modified state \end{small}
+
 \end{algorithm}
@@
 In that process, tuning correctly the $\gamma$ parameter is important.
-In Figure~\ref{fig:discounts}, we represent a set of actions that to a gain - $s(\B{x^*} = 1$ here, and the discount function that we apply is $\gamma^{\Delta-\delta}s(\B{x^*})$.
+In Figure~\ref{fig:discounts}, we represent a set of actions that to a gain - $s(\B{x_c} = 1$ here, and the discount function that we apply is $\gamma^{\Delta-\delta}s(\B{x_c})$.
 
@@
 \begin{center}
-\includegraphics[scale=.5]{"discounts"}
+\includegraphics[scale=.4]{"discounts"}
 \caption{Different discount rates affecting intermediary states}
@@
 
-\subsection{Convergent estimates for equivalent underlying states}
+We are going to develop a simple insight about why and how the proposed algorithm might lead to optimal long term decisions.
+For this, we recall the important steps of the algorithm.
+Firstly, it evaluates the best possible action to take after observing a stochastic state $\B{x_t}(p_t)$.
+Secondly, it appends the latter or a randomly chosen one to the training features matrix $\B{X}$.
+If the signal function $s(\B{x^*})$ is triggered, it backwardly assigns values to the states of the episode thanks to the discount function $d(\gamma, \Delta, \delta, s(\B{x^*})$.
+These values are finally appended to the training target variable $\B{y}$.
+Then the training set is used as an input for the approximation function $f()$ to find $\B{w} = \arg\min_{\B{w}}(l(f(\B{w},\B{X}),\B{y}))$.
+In the next episode, the algorithm evaluates a new stochastic state and uses the new $\B{w}$ to make take the next decision.
 
-The convergence of the algorithm's evaluations is closely linked to the performance of the regressor function at extracting underlying patterns between states and values, assuming however that this relationship exists.
+An important assumption has to be satisfied for the algorithm to take the right decisions.
+There needs to be patterns in the stochastic states.
+If it is not the case, the approximation function is unable to optimise the value of $\B{w}$ with time.
+But it does not make sense to use such algorithm if we know that we are in such setting.
+By opposition, an application of choice is when we have imperfect representations of the same underlying states, such has pictures of a board game.
 
-Let's consider a finite set of underlying states.
-If the conditions are favorable, we expect overtime, as the training sample is fed after each episode, a better evaluation of the different states and, asymptotically, underlying equivalent states will get the same value.
-Under another perspective, we expect $w^*$ to converge as $X$ grows.
+Another assumption is that $s(\B{x^*})$ is able to give an objective information about the current state.
+This is a strong assumption.
+However, we can imagine a practical solution for this.
+This signal function could result from a pre-trained supervised approximation function that was able to evaluate with high precision the features’ values triggering each of the possible signal.
+Another solution, providing perfect information, is manual monitoring at the end of each episode.
 
-An interesting point is the expected non-linearity of this process.
-Indeed, as long as the learner did not discover all gain-yielding states, the regressor's parameters do not permanently stabilize.
-Once all states have been uncovered, this convergence happens.
+Finally, we require a correct tuning of the algorithm’s different elements.
+The approximation function should be complex enough to recognise the possible patterns in $\B{x_t}(p_t)$.
+$\Omega(\B{x_t})$ should be well designed.
+The discount function $d(\gamma, \Delta, \delta, s(\B{x^*}))$ should also be well designed.
 
-If the set of underlying states is infinite, the regressor's parameters can stabilize only if there exist patterns between the states and the value.
-An example of this situation is a vision-equipped learner whose task is to catch the same ball in different environments.
-
-\subsection{Long term gain maximization}
+With these conditions fulfilled, we firstly expect that, as the training sample is fed after each episode, a better evaluation of the different states and, asymptotically, underlying equivalent states will get the same value.
+Formally, we anticipate:
+$$\B{w} \xrightarrow[n \to \infty]{} \B{w^*}$$
+where $n$ is the number of previously observed states and $w^*$ is the optimal set of weights. In most cases, we expect the convergence process to be non-linear.
 
 Evaluating well the value of equivalent states is only a first step to obtain a learner that is useful in practice.
-To behave in a long term-maximizing way, the algorithm needs to have a relatively balanced experience between gains and losses.
+To maximise the long-term gains, the algorithm needs to have a relatively balanced experience between gains and losses.
+First, the tuning of $\gamma$ should allow the regressor to evaluate states faster, via the learner's actions.
+If necessary, we need to insure that
+$d(\gamma, \Delta, \delta, s_{gain}) \simeq d(\gamma, \Delta, \delta, s_{loss})$
+if a $\delta$th intermediary action possibly leads to a gain in one episode but also to a loss in another episode.
+Second, tuning $\epsilon$ well will insure that the learner explores potentially good actions fast enough ; an episode number-dependent value for $\epsilon$ might be desirable.
+If these final assumptions are insured, we expect:
+$$\B{x_c} \xrightarrow[n \to \infty]{} \B{x^*}$$
+where $x^*$ is the optimal action for a particular time $t$.
 
-First, tuning $\gamma$ in a proper way will help to allow the regressor to evaluate states well faster, via the learner's actions. If possible, a good way to tune this parameter is to make such that
-$$d(\gamma, \Delta, \delta, s_{gain}) \simeq d(\gamma, \Delta, \delta, s_{loss})$$
-for the $\delta$th intermediary action that led to a gain in one episode but also to a loss in another episode.
+\section{Implementation for a tic-tac-toe robot player}
 
-Second, tuning $\epsilon$ well will insure that the learner explores potentially good actions fast enough. A number of episodes-dependent $\epsilon$ might be desirable in that regard.
+Tic-tac-toe is a two players game where the players are represented by the symbols X and O in a three by three grid.
+Each player takes turns marking the spaces with their symbol.
+The player who places three of his symbol in a horizontal, vertical, or diagonal row wins the game.
+Note that draw is a possible outcome.
 
-\section{Implementaion for a "real-world" tic-tac-toe game}
+Robots playing this game are a classical application in RL.
+But this time we want the learner to play according to its \textit{vision} of a board that sequentially drawn with a pen.
+The information available to the robot corresponds to images transformed into vectors.
+Ultimately, the learner has to discover that the images shown in Figure~\ref{fig:gameA} represent the same game.
 
-The tic-tac-toe game is a game with initially nine positions organised as a three by three square.
-After the first player played the first move, the second player is left with eight available moves.
-This is until one player won or once the whole board is covered, yielding a draw.
-To win, a player must align his moves on the same line or on the same column, or by filling one of the board's diagonal.
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.14]{"gameA"}
+\includegraphics[scale=.35]{"gameA_"}
+\caption{After enough games, the learner is able to play the same way after visualizing either of these images}
+\label{fig:gameA}
+\end{center}
+\end{figure}
 
-The tic-tac-toe game is a RL classical application, but this time the learner has to play according to his "vision".
-Notably, when playing against a human, the information available to the robot corresponds to an image of the type:
+The images of the board are thus stochastic representations of a finite set of underlying states.
+The stochastic processes affecting particularly these representations are the variations in the handwriting and the brightness of the environment.
 
-TWO IMAGES
-caption: after enough games, the learner should be able to play the same way after visualizing either of these images
-
-This is thus a "stochastic representation" from a finite set of states. The stochastic processes affecting this representation are notably the varying handwrittings and the environment's brightness.
-
-After a sufficient number of games, the robot is able to effectively assign values to the potential actions it might take:
-ADD 150*150 PICTURE WITH PROBAS ON TOP, FOR 100, 1000, 5000 selfplays
+After a sufficient number of games, the robot is able to effectively assign values to the potential actions it might take, as displayed in Figure~.
+ADD 150*150 PICTURE WITH PROBAS ON TOP, FOR 100, 1000, 5000 self plays
 REVALUE SO LOSS = -1, DRAW=0
 
-\subsection{Particularities of the implementation}
+\subsection{Particularities of the algorithm's implementation}
 
-The chosen regressor is an artificial neural network detailed in Figure~\ref{fig:neuralnet}.
-I chose this type of regressor as I believe it supports the complex relationship between the pixels representing a state and the value of this state.
-A sufficiently complex structure is thus able to handle the very large number of ways to represent a game combined with the many game combinations.
-In this implementation, the network is has the following form:
+The general algorithm presented in the previous section had to be adapted for this particular application, in particular regarding the approximation function, the signal function and the discount function.
+
+The chosen approximation function is an artificial neural network detailed in Figure~\ref{fig:neuralnet}.
+It outputs real values and is a regressor.
+The choice for this regressor was driven by the complexity of the relationship existing between the representations of the states and the value assigned to these states.
+Indeed, this particular architecture proved to capture the infinite number of possible representations of $255168$ possible states, without taking into account rotations and reflections.
 
@@
 
-The stochastic state $x_t$ is resulting from another player's move represented on an image as seen above.
-In terms of strategy optimization, here the learner would have to think ahead.
-The states here are the fruit of *rival* actions, either from a human player or from another robot.
-And in this implmentation, the learner plays after the rival, so it can win either in three or four moves.
+The stochastic state $\B{x_t}(p_t)$ is resulting from another player's move.
+The states are the fruit of \textit{rival} actions, either from a human player or from another robot.
+So in terms of strategy optimization, the learner has to anticipate a possibly defavorable environment during the game and act accordingly.
+In particular, in this implementation, the learner plays after the rival, so it can win only with three or four moves.
 
-The signal function is set by a human.
-It triggers the values $1$ if the learner wins, $-0.1$ if there is a draw and $-1$ if it loses or if it does an illegal move.
+The signal function is set manually.
+It is triggered when a game ends.
+If the learner wins, it triggers $1$, $-0.1$ if there is a draw and $-1$ if it loses or if it does an illegal move.
 Notice that an illegal move corresponds only to playing on top of the other player.
@@
 
-In this implementation, the discount function $d(\gamma, \Delta, \delta, s(x^*))$ has the form
-$$\gamma^{\Delta-\delta}s(x^*)$$
-More precisely, I picked $\gamma=0.3$ as it yields the same value for an early move.
-The discounted value for moves leading to a win thus are: 1, 0.3, 0.09 and 0.081.
-For a draw: -0.1, -0.01, -0.001 and -0.0001.
-For a loss: -1, -0.3, -0.09 and -0.081.
-For an illegal move, only the signal value is given to the last move.
+In this implementation, the discount function $d(\gamma, \Delta, \delta, s(\B{x_c}))$ has the form
+$$\gamma^{\Delta-\delta}s(\B{x_c})$$
+More precisely, I picked $\gamma=0.3$ so we get neutral values, close to $0$, for early moves regardless of the game's outcome.
+The possible discounted values for moves leading to a win thus are: $1$, $0.3$, $0.09$ and $0.081$.
+For a draw: $-0.1$, $-0.01$, $-0.001$ and $-0.0001$.
+For a loss: $-1$, $-0.3$, $-0.09$ and $-0.081$.
+When an illegal move happens, there is no discount computation and this move gets the value $-1$.
 
-I implemented an adaptive randomness factor. It has the very *ad hoc* form 
-$$\epsilon = 50000/log_{10}(S)^6,$$
+Finally, I implemented an adaptive randomness factor. It has the very \textit{ad hoc} form 
+$$\epsilon = 30000/log_{10}(S)^6,$$
 where $S$ is the size of the data in bytes.
-I chose this formula so it equals 99\% for a data corresponding to 10000 games.
-The other learner is penalized and has $\epsilon = 100000/log_{10}(S)^6$ so our learner has time to learn how to win.
+I chose this formula so it equals 99\% for a data corresponding to 10000 games which is how many games I train the learner with.
+The other robot is penalized with $\epsilon = 120000/log_{10}(S)^6$ so our learner has time to learn how to win in the beginning.
+This parameter has to be set sensibly for an optimized training to get a perfect "player".
+However we will not study this aspect further here.
 
@@
 After each of the players' move, an image representing the current state of the board is generated.
-This is image is then reduced and appended to the training set.
-SHOW LARGE IMAGE -> SMALL IMAGE transition
+This is image is then reduced and appended to the training set, as shown in Figure~\ref{fig:transition}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.4]{"discounts"}
+\caption{Different discount rates affecting intermediary states}
+\label{fig:transition}
+\end{center}
+\end{figure}
 
@@
 
-INCLUDE ADAPTATIVE LEARNING RATE?
-
 Notice that a learner trained against a random opponent is able, after 100 games, to play easy moves, typically the last move to win, against a human player.
@@
 The learner can only win in three or four moves. Against a random player, we that the learner has an average number of moves to win closer to three with more games.
-
-\subsection{Improvements}
-
-Convolutional layer
-
-Detect all illegal moves
 
@@
 
+\bibitem{c6} Munos et al. (2016), Safe and efficient off-policy reinforcement learning, arXiv:1606.02647v1 [cs.LG]
+
+\bibitem{c7} Seijen et al. (2015), True Online Temporal-Difference Learning, arXiv:1512.04087v1 [cs.AI]
+
 \end{thebibliography}

### 2016-06-25 tic21.py

@@ -107,16 +107,18 @@ class Learner:
 
         s.net = Regressor(
                 layers=[
-                    Convolution("Rectifier", channels=81, kernel_shape=(3,3), pool_shape=(2,2)),
-                    Layer("Sigmoid", units=200),    
-                    Layer("Rectifier", units=200),
+                    Convolution("Rectifier", channels=27, kernel_shape=(3,3), kernel_stride=(2,2)),
+                    Convolution("Rectifier", channels=54, kernel_shape=(2,2), kernel_stride=(1,1)),
+                    #Convolution("Rectifier", channels=27, kernel_shape=(3,3), kernel_stride=(1,1)),
+                    Layer("Sigmoid", units=210),
+                    Layer("Rectifier", units=210),
                     Layer("Sigmoid")],
-                n_iter=50,
-                learning_rate=0.0001,
+                n_iter=100,
+                learning_rate=0.001,
                 parameters=params,
                 valid_set=cv,
                 learning_rule='rmsprop',
-                f_stable=0.0001,
+                f_stable=0.001,
                 verbose=True,
                 batch_size=200,
                 n_stable=10)

### 2016-06-28 paper2.tex
yo

@@
-\documentclass[letterpaper, 10.5 pt, conference]{ieeeconf}
+\documentclass[scrartcl, 10.5 pt, conference]{ieeeconf}
 \IEEEoverridecommandlockouts
@@
 \title{\LARGE \bf
-Automated consistent course of actions in a stochastic environment
+Dynamic actions in stochastic environments \\
+\large \& application with neural network approximations-based reinforcement learning
 }
@@
 
-I propose an algorithm allowing the automated learning of the best course of actions in a stochastic environment.
+I propose an algorithm allowing an artificial entity to optimize its sequence of actions in a stochastic environment.
 It notably includes a signal function that should allow more flexible implementations than current approximation-based reinforcement learning algorithms.
-This function triggers the computation of discounted action values for each learning episode.
-The paper contains an application, a tic-tac-toe robot player recognizing underlying games from photographs via a neural network.
+In the fifth section, I present an implementation of the algorithm, it is a tic-tac-toe robot player recognizing underlying games from photographs thanks to pattern recognition via a convolutional neural network.
 
@@
 This entity needs two abilities: learning from its own actions and recognizing useful patterns in the current state.
-The first ability typically corresponds to algorithms from the reinforcement learning (RL) field.
-The second ability requires regressor or classifier function that are able to handle the complexity of stochastic state's patterns.
+The first ability typically corresponds to reinforcement learning (RL) algorithms.
+The second ability requires a regressor or a classifier function that is able to handle the complexity of stochastic state's patterns, provided that they exist.
 
@@
 RL emerges easily from other Machine Learning techniques as the training data is generated thanks to the actions of the learning entity.
-Thus, the \textit{learner} uses training information that evaluates the actions taken rather than instructs by giving correct actions are faced repeatedly with a choice among $n$ different options, or actions.
-Such setting has its roots in the \textit{n-armed bandit problem} where after each choice you receive a numerical reward chosen from a stationary probability distribution that depends on the selected action.
-The objective is to maximize the expected total reward over all the episodes.
-The learner evolves in a particular state and records what actions lead to a maximal gain.
+Thus, the \textit{learner} uses its history of actions to evaluate what is the best move to do next.
+It needs for this to have faced repeatedly identical or similar situations.
+Such setting has its roots in the \textit{n-armed bandit problem} where after each choice a numerical reward is received which is chosen from a stationary probability distribution that depends on the selected action.
+The learner records what actions led to a maximal gain.
 In most cases the goal is to maximize a long term gain by taking a correct course of actions.
-These policies can be more or less explicit.
-I our case, the proposed algorithm would be qualified as \textit{off-policy}.
 
-An RL learner is thus particularly dependent on a state signal that succeeds in retaining all relevant information, a \textit{Markov} state.
-For example, a checkers position (the current configuration of all the pieces on the board) would serve as a Markov state because it summarizes everything important about the complete sequence of positions that led to it.
+An RL learner is thus particularly dependent on a state that retains all relevant information, a \textit{Markov} state.
+For example, a checkers position - the current configuration of all the pieces on the board - would serve as a Markov state because it summarizes everything important about the complete sequence of positions that led to it.
 Much of the information about the sequence is lost, but all that really matters for the future of the game is retained.
@@
 A reinforcement learning task that satisfies the Markov property is called a Markov Decision Process (MDP).
-If the state and action spaces are finite, it is possible the MDP framework notably formalizes the probability that a state occurs given a previous state an action, and the probability of reward after the new state occurred.
+If the state and action spaces are finite, the MDP framework notably formalizes the probability that a state occurs given a previous state an action, and the probability of reward after the new state occurred.
 
@@
 
-Equipping the learner with an approximation function enables it to evaluate an infinite set of states but also a finite set of underlying states having an infinite set of representations.
+Equipping the learner with an approximation function can enable it to evaluate an infinite set of states but also a finite set of underlying states having an infinite set of representations.
 In Figure~\ref{fig:surf}, I illustrate the fact that humans are able to recognize the same underlying situation and take the correct course of actions.
@@
 
-In a similar manner, we would like our learner to overcome the imperfect information it receives and learn how to behave in the long term to maximize its gain.
-This is made possible notably by regressor and classifier functions that take particularly well advantage of naturally generated data such as images and sounds.
+The \textit{imperfect representations} case is the one that interest us here.
+We thus would like our learner to overcome the imperfect information it receives and learn how to behave consistently during a whole episode to maximize its gain.
+This requires the use of regression or classification functions that are can process efficiently patterns in the different states.
 
-A modified version of the classical tic-tac-toe robot will provide a good example for the principle we seek to study.
+In section V., a modified version of the classical tic-tac-toe robot will provide a good example for this principle.
+It required a regression function that could take particularly well advantage of the structure of image data.
 
@@
 
-The proposed algorithm posseses similarities with the concepts that we are presenting in this section.
-We also are going to show that, to my knowledge, it is a novel approach in approximation-based RL.
-This would notably be due to the algorithm's minimal requirements for control of the states' values.
+The proposed algorithm possesses similarities with the concepts presented in this section.
+I also am going to show that, to my knowledge, it is a novel approach in approximation-based RL ; this would notably be due to the algorithm's minimal requirements for the monitoring of the states' values.
 
-A close classical RL technique to the proposed algorithm, is the the State-Action-Reward-State-Action algorithm (SARSA).
-It uses approximations to update the value of a whole set of actions, a policy.
+A classical RL technique that is close to the proposed algorithm, is the State-Action-Reward-State-Action algorithm (SARSA).
+It uses approximations to update the value of a whole set of actions called a \textit{policy}.
 SARSA is well suited when the value of each action can be monitored.
-It notably update the parameters of the approximation function after each movement thanks to a change in value from $t-1$ to $t$, known as a temporal difference.
-The update includes a gradient term, compute from the difference between the state's value approximation and the real state's value.
-Our algorithm however does not suppose the learning of a whole policy.
+It notably updates the parameters of the approximation function after each movement thanks to a change in value from $t-1$ to $t$, known as a temporal difference.
+The update includes a gradient term, computed from the difference between the state's value approximation and the real state's value.
+The proposed algorithm is fundamentally close to SARSA, but it not suppose the learning of a whole policy.
 
 In that sense, the recent Deep Q networks (DQN) are closer to what is proposed.
-This off-policy algorithm uses neural network approximations of the states' value to decide on an action.
+This \textit{off-policy} algorithm uses neural network approximations of the states' value to decide on an action.
 And as the learner makes better moves, the neural network is fed with more useful information.
-And thanks to its \textit{experience replay}, when training the network, random mini-batches from the replay memory are used instead of the most recent transition.
+And thanks to its \textit{experience replay} feature, when training the network, random mini-batches from the replay memory are used instead of the most recent transition.
 Such efficient implementation makes the DQN very adaptive to different tasks, as demonstrated for the learning of multiple arcade games while using the same network architecture.
-I also assume such memory device for the proposed algorithm.
+I also implemented such memory device for the application of the proposed algorithm.
 But I try to have a more flexible approach than this algorithm, which supposes the constant monitoring of the states' values.
 
-This is done thanks to a mechanism that is close to the True Online Temporal Difference Learning from Seijen et al..
-The authors introduced a formal way to assign values in a backward manner to each state of the episode, when the latter ends.
-This is done in a discounted way so states farther to the end of an episode get a value with a smaller magnitude.
+This is done thanks to a mechanism that is related to the True Online Temporal Difference Learning from Seijen et al..
+The authors introduced a formal way to update the values in a backward manner for each state of the episode, when the latter ends.
+This is done in a discounted way so states farther from the end of an episode get a value with a smaller magnitude.
 In the proposed algorithm, however, as mentioned earlier, we want to avoid using a permanent control of the states' values during the learning phase.
@@
 
-Hopefully the proposed algorithm offers a greater flexibility than the previous litterature.
+Hopefully the proposed algorithm offers a better flexibility those from the previous literature.
 Removing the assumption that the state's values are permanently monitored during the training phase is made possible by the implementation of a \textit{signal function} coupled with a \textit{discount function}.
@@
 When triggered, it assigns a gain or loss-related value to the final state of the episode.
-The second function will allow to assign values to each state of a particular episode.
-This mechanism is supposed to allow a practical use of the algorithm when evaluating non-idealized states - not only video games screen shots - involving stochastic perturbations that are for instance recorded via a visual sensor.
+The second function will allow to assign values to each state of an episode.
+This mechanism is supposed to allow a practical use of the algorithm when evaluating non-idealized states - thus not only video games screen shots - involving stochastic perturbations that are for instance recorded via a visual sensor.
 
@@
 
-We can summarise our setting as a decision process possibly lacking the Markov property. It has the form
+We can summarize our setting as an MDP having the form
 $$\langle \mathcal{X},\Omega, P(\B{x_c},\B{x_{t+1}}), \B{y},s(\B{x_c}),d(\gamma, \Delta, \delta, s(\B{x_c})) \rangle$$
-where $\mathcal{X}$ is a set of finite underlying states, $\Omega$ is the finite set of possible actions, $P(\B{x_c})$ is the probability that the process moves to a new state $\B{x_{t+1}}$ potentially depending on numerous previous states in the episode, $s()$ is the signal function and d() is a the discount function.
+where $\mathcal{X}$ is the set of finite underlying states, $\Omega$ is the finite set of possible actions, $P(\B{x_c})$ is the probability that the process moves to a new state $\B{x_{t+1}}$ depending on the current state and the decision of a particular action, $s()$ is the signal function and $d()$ is the discount function.
 
-I am going to describe more details and chronologically what happens during an episode of the algorithm in action.
+I am going to describe chronologically what happens during an episode of the algorithm in action.
 Note that all the objects containing $X$ or $x$ in their names are $m$-dimensional objects, where $m$ is the number of features representing the considered states.
 
-The learner observes a new stochastic state $\B{x_t}(p_t)$ - note that this stochastic state can be affected by previous moves from our learner.
+The learner observes a new stochastic state $\B{x_t}(p_t)$ - note that this stochastic state can be affected by previous moves from the learner.
 It then simulates actions within the finite set of possible actions $\Omega(\B{x_t})$.
 $\Omega(\B{x_t})$ can be set deterministically.
-For instance, in the context of a game, there is a well determined set of actions that a human can tune in advance.
-It also can be a subset of an infinite set of actions.
-For instance, a mobile robot can explore the real 3D world in almost infinite ways.
-In that case, the set of possible actions has to be restricted somehow, possibly by the time allowed to go in particular direction.
+For instance, in the context of a game, there is a well determined set of actions that can be tuned in advance.
+In the case of a mobile robot physically exploring the real 3D world, the set of possible actions has to be restricted somehow, possibly by the time allowed to go in a particular direction.
 
-All of these simulated actions are stored in a temporary matrix $\B{X'}$.
-Each observation in this matrix is going to be evaluated by the a approximation function $f(\B{w},\B{X'})$.
+All the simulated actions are stored in a temporary matrix $\B{X'}$.
+Each observation in this matrix is evaluated by the a approximation function $f(\B{w},\B{X'})$.
 This function can be a regressor or a classifier which weights $w$ are randomly initialized before the first episode.
-It is going to return approximations of the current state's value.
-This value is proportional to the learner's closeness to its objective.
-Among the simulated states, we will then return the chosen $\B{x}$, $\B{x_c}$.
-With probability $1-\epsilon$ it corresponds to the state which has the highest approximated valu and with probability $\epsilon$, it will be a randomly chosen state from $\B{X'}$.
+It then approximates each simulated state's value.
+Among the simulated states, the chosen state, $\B{x_c}$ will be, with probability $1-\epsilon$ the one which has the highest approximated value.
+And with probability $\epsilon$, it will be a randomly chosen state from $\B{X'}$.
 If both cases, we append the returned state to the training set $\B{X}$.
@@
 The learner possesses a function $s(\B{x_c})$.
-It returns different values when a particular signal in the state generated by the learner's last move, $\B{x_c}$, is sensed.
-The value is positive if a \textit{gain} is sensed, and negative value if a \textit{loss} is sensed.
-At least one value should be potentially triggered.
-A typical example for $[gain;loss]$ would be respectively $[1;-1]$.
+It returns different values when a particular signal, function of the state generated by the learner's last move, is triggered.
+Its value is positive if a \textit{gain} is sensed, and negative value if a \textit{loss} is sensed.
+The signal function should at return one possible value.
+A typical example for $\{gain;loss\}$ would be respectively $\{1;-1\}$.
 While this signal is off, the learner counts the number of actions since the signal was last triggered.
 When the signal is triggered, the current episode is over.
-Each action of the episode receives a value given by the discount function $d(\gamma, \Delta, \delta, s(\B{x_c}))$ where $\gamma$ is a discount rate, $\Delta$ is the number of actions in the episode and $\delta$ is the index of the action in the episode.
-For instance, the last action in the episode has $\delta=\Delta$.
-The function should insure that actions that are closer in time from the triggered signal get values that are closer from $s(\B{x_c})$.
+Each action of the episode receives a value given by the discount function $d(\gamma, \Delta, \delta, s(\B{x_c}))$ where $\gamma$ is a discount rate, $\Delta$ is the number of actions in the episode and $\delta$ is the index of the action in the episode - the last action in the episode thus has $\delta=\Delta$.
+The discount function should insure that actions that are closer in time from the triggered signal get values that are closer from $s(\B{x_c})$.
 Each value of the episode are then appended to the training target vector $\B{y}$.
 
-We then get a training sample on which are going to apply our approximation function $f(\B{w}, \B{X})$ which this time minimizes the loss $l(f(\B{w}, \B{X}),\B{y})$.
+We then get a training sample on which are going to apply our approximation function $f(\B{w}, \B{X})$ which aim is to minimize the loss $l(f(\B{w}, \B{X}),\B{y})$.
 From this process, we are going to retain set of parameters corresponding to the minimal loss to update $w$.
-The updated $w$ is going to be used in the next episode in order to predict the value of simulated actions.
+The updated $w$ is going to be used in the next episode to predict the value of simulated actions.
 
-Algorithm~\ref{alg:algo} is a formal version of the proposed algorithm.
+Below is presented a formal version of the proposed algorithm.
 
@@
 \vspace{3mm}
-\begin{small} $m$: number of features, $\B{X}=\B{X_{\o}}$: empty $m$-columns matrix, $\B{y}=\B{y_{\o}}$: empty target vector, $\B{w}=\B{w_0}$: random weights for the regressor, $\gamma \in [0,1]$  discount rate, $\Delta = 0$: number of actions, $f()$: regressor function, $p_t(\B{x_c})$: stochastic process, function of the last existing chosen state, $\B{X'}$: $m$-columns simulated states matrix, $\B{x_t}(p_t) \in {\rm I\!R^m}$: current state vector, $\Omega_t(\B{x_t})$: set of possible actions, $\epsilon$: randomness rate, $s(\B{x_c}) \in {\rm I\!R}$: signal value potentially triggered by the modified state \end{small}
+\begin{small} $m$: number of features, $\B{X}=\B{X_{\o}}$: empty $m$-columns matrix, $\B{y}=\B{y_{\o}}$: empty target vector, $\B{w}=\B{w_0}$: random weights for the approximation function, $\gamma \in [0,1]$  discount rate, $\Delta = 0$: number of actions, $f()$: approximation function, $p_t(\B{x_c})$: stochastic process, function of the last existing chosen state, $\B{X'}$: $m$-columns simulated states matrix, $\B{x_t}(p_t) \in {\rm I\!R^m}$: current state vector, $\Omega_t(\B{x_t})$: set of possible actions, $\epsilon$: randomness rate, $s(\B{x_c}) \in {\rm I\!R}$: signal value potentially triggered by the modified state \end{small}
 
@@
 
-\subsubsection{Regressor function}
+\subsubsection{Approximation function}
 
-One interesting aspect is the role of the regressor function.
-Contrarily to typical applications, the functions's task is utimately not to predict an action's value with maximum accuracy, even though it is often related to what we want, but rather to give the "real value" to an action.
-Imagine for instance a game where two intermediary states might lead to both winning and losing.
-Such states will get a non-tendential value from the function, which is a good thing.
+One interesting aspect is the role of the regression function.
+Contrarily to typical applications, the function's task is ultimately not to predict an action's value with maximum accuracy, even though it is often related to our objective, but rather to give the "real value" to an action.
+Imagine for instance a game where an intermediary state leads to both a gain or a loss.
+With enough iterations, such states should get a non-tendential value from the function.
 
-Even if the approximation function's task is this time to unveil the real value of each state, the choice for the right function is still clearly very important.
-It should notably depend on the complexity of the patterns that we wish the learner to discover, which is often proportional to the complexity of the considered $m$-dimensional states.
-
-Obviously, if these states approximately form a linear function of the target value, a linear approximation function can be applied.
-But if the states are strongly affected by stochastic processes, other functions could be considered.
+It is crucial to make the choice of the approximation function based on the task in question.
+What matters here is the type of the patterns that we wish the learner to discover and based on these patterns' complexity, which is often proportional to $m$ dimension of the states.
+Then, if these states are approximately linearly related to the target value, a linear approximation function can be applied.
+But if the states are strongly affected by stochastic processes, other functions should be considered.
 
 Typically in this paper, we consider real-world representations as states.
-In other words, our input could be include images, sounds and texts.
+In other words, our input could be images, sounds or texts.
 In that case, we might consider using neural networks which are often celebrated as a particularly accurate family of algorithms for such data.
-Using neural networks, we can see our learner as an imitator of how humans would act when facing the task in question.
-The challenge is then to recognize equivalent underlying states thanks to pattern recognition.
-Finally note, considering the algorithm described above, that $\B{w}$ would be the set of a neural network's hidden layers.
+Using neural networks, we can see our learner as imitating how humans would act when facing specific tasks.
+The challenge is then to recognize equivalent real-world states thanks to pattern recognition.
+Considering the algorithm described above, $\B{w}$ would in that case be the set of a neural network's hidden layers.
 
@@
 The signal function allows to reduce the assumptions on the basic abilities of the learner.
-While most approximation-based reinforcement learning algorithms suppose the permanent presence of a value-monitoring function, here we only assume the presence of a discrete, time-independent sensor triggering one or more values.
+While most approximation-based reinforcement learning algorithms suppose the permanent presence of a value-monitoring function, here we only assume the presence of discrete, time-independent signals that can have one or more values.
 
-The use of discounted values enables each step to get a value at the end of each episodes.
+The use of discounted values enables each state to get a value at the end of each episode.
 In that process, tuning correctly the $\gamma$ parameter is important.
@@
 We see that with a larger discount factor, more emphasis is put on the final result of the set of actions.
-If relevant, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.
+If relevant, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and to a loss.
 
@@
 But the learner still looks for best discounted value according to the current state.
-Thus, with less assumptions on the monitoring of the real value of the states, we still benefit from something common to RL algorithms: the discovery and use of intermediary objectives.
+Thus, with less assumptions on the monitoring of the real value of the states, we still benefit from something common to RL algorithms, the discovery and use of intermediary objectives.
 
-\section{Behaviour of the algorithm}
+\section{Behavior of the algorithm}
 
-We are going to develop a simple insight about why and how the proposed algorithm might lead to optimal long term decisions.
-For this, we recall the important steps of the algorithm.
+Here is developed a simple insight on why and how the proposed algorithm might lead to optimal long term decisions.
+For this, we summarize the important steps of the algorithm.
 Firstly, it evaluates the best possible action to take after observing a stochastic state $\B{x_t}(p_t)$.
-Secondly, it appends the latter or a randomly chosen one to the training features matrix $\B{X}$.
+Secondly, it appends the latter or a randomly chosen possible action to the training features matrix $\B{X}$.
 If the signal function $s(\B{x^*})$ is triggered, it backwardly assigns values to the states of the episode thanks to the discount function $d(\gamma, \Delta, \delta, s(\B{x^*})$.
@@
 
-An important assumption has to be satisfied for the algorithm to take the right decisions.
-There needs to be patterns in the stochastic states.
-If it is not the case, the approximation function is unable to optimise the value of $\B{w}$ with time.
-But it does not make sense to use such algorithm if we know that we are in such setting.
-By opposition, an application of choice is when we have imperfect representations of the same underlying states, such has pictures of a board game.
+An important assumption has to be satisfied for the algorithm to function correctly: the presence of patterns in the stochastic states.
+Without patterns, the approximation function is unable to optimize the value of $\B{w}$ with time.
+That said, it does not make sense to use the algorithm in such setting.
+By opposition, an application of choice is when we have imperfect representations of the same underlying states, such has real-world representations of a particular task.
 
-Another assumption is that $s(\B{x^*})$ is able to give an objective information about the current state.
+We also assume that $s(\B{x^*})$ is able to give an objective information about the current state.
 This is a strong assumption.
 However, we can imagine a practical solution for this.
-This signal function could result from a pre-trained supervised approximation function that was able to evaluate with high precision the features’ values triggering each of the possible signal.
-Another solution, providing perfect information, is manual monitoring at the end of each episode.
+This signal function could result from a pre-trained supervised approximation function that is able to evaluate with high precision the features’ values triggering each of the possible signals.
+Another solution, providing perfect information, is a manual monitoring at the end of each episode.
 
 Finally, we require a correct tuning of the algorithm’s different elements.
-The approximation function should be complex enough to recognise the possible patterns in $\B{x_t}(p_t)$.
+The approximation function should be complex enough to recognize the possible patterns in $\B{x_t}(p_t)$.
 $\Omega(\B{x_t})$ should be well designed.
@@
 
-With these conditions fulfilled, we firstly expect that, as the training sample is fed after each episode, a better evaluation of the different states and, asymptotically, underlying equivalent states will get the same value.
+With these conditions fulfilled, we firstly expect, as the training sample is fed after each episode, a better evaluation of the different states.
+Secondly, we expect that, asymptotically, underlying equivalent states will get the same value.
 Formally, we anticipate:
@@
 
-Evaluating well the value of equivalent states is only a first step to obtain a learner that is useful in practice.
-To maximise the long-term gains, the algorithm needs to have a relatively balanced experience between gains and losses.
-First, the tuning of $\gamma$ should allow the regressor to evaluate states faster, via the learner's actions.
+Evaluating well the value of equivalent states is only a first step to obtain a useful learner in practice.
+To maximize the long-term gains, the algorithm needs to have a relatively balanced experience between gains and losses.
+First, the tuning of $\gamma$ should allow the regression function to evaluate states faster, via the learner's actions.
 If necessary, we need to insure that
@@
 The player who places three of his symbol in a horizontal, vertical, or diagonal row wins the game.
-Note that draw is a possible outcome.
+A draw is a possible outcome.
 
 Robots playing this game are a classical application in RL.
-But this time we want the learner to play according to its \textit{vision} of a board that sequentially drawn with a pen.
+But this time we want the learner to play according to what is seen on a real-world image.
+The learner will then evaluate game boards that are sequentially drawn with a pen.
 The information available to the robot corresponds to images transformed into vectors.
@@
 The images of the board are thus stochastic representations of a finite set of underlying states.
-The stochastic processes affecting particularly these representations are the variations in the handwriting and the brightness of the environment.
-
-After a sufficient number of games, the robot is able to effectively assign values to the potential actions it might take, as displayed in Figure~.
-ADD 150*150 PICTURE WITH PROBAS ON TOP, FOR 100, 1000, 5000 self plays
-REVALUE SO LOSS = -1, DRAW=0
+The stochastic processes affecting particularly these representations are the variations in the handwriting and the brightness level of the images.
 
@@
 
-The general algorithm presented in the previous section had to be adapted for this particular application, in particular regarding the approximation function, the signal function and the discount function.
+The general algorithm presented in the previous section had to be adapted for this application, in particular regarding the approximation function, the signal function and the discount function.
 
-The chosen approximation function is an artificial neural network detailed in Figure~\ref{fig:neuralnet}.
-It outputs real values and is a regressor.
-The choice for this regressor was driven by the complexity of the relationship existing between the representations of the states and the value assigned to these states.
-Indeed, this particular architecture proved to capture the infinite number of possible representations of $255168$ possible states, without taking into account rotations and reflections.
+First note that the stochastic state $\B{x_t}(p_t)$ is resulting from another player's move.
+The states are the fruit of \textit{rival} actions, either from a human player or from another robot.
+So in terms of strategy optimization, the learner has to anticipate a possibly unfavorable environment during the game and act accordingly.
+In particular, in this implementation, the learner plays after the rival, so it can win only with three or four moves.
+
+The chosen approximation function is a convolutional neural network detailed in Figure~\ref{fig:neuralnet}.
+This regression function outputs real values.
+The choice for this regression function was driven by the complexity of the relationship existing between the representations of the states and the value assigned to these states.
+The convolutional layer greatly enhanced the precision of the approximations with this training set formed by vectorized images.
+Indeed, this architecture proved to capture well patterns in the infinite number of possible representations of $255168$ possible ways to finish the game.
 
@@
 \begin{center}
-\includegraphics[scale=.25]{"NNview"}
-\caption{Other parameters are: the number of iterations = 200, the learning rate = 0.0001, the learning rule = Rmsprop which divides the learning rate for a weight by a running average of the magnitudes of recent gradients for that weight, the threshold under which the validation error change is assumed to be stable = 0.0001, the mini-batch size = 200 observations, and the number of iterations after which training should return when the validation error remains (near) constant = 10.}
+\includegraphics[scale=.3]{"NNview"}
+\caption{Additionally, the number of iterations is 200, the learning rate is 0.0001, the learning rule is Rmsprop which divides the learning rate for a weight by a running average of the magnitudes of recent gradients for that weight, the threshold under which the validation error change is assumed to be stable is 0.0001, the mini-batch size is of 200 observations, and the number of iterations after which training should return when the validation error remains (near) constant is 10.
+Note that the output is then transformed so it belongs to the range $[-1;1]$.}
 \label{fig:neuralnet}
@@
 
-The stochastic state $\B{x_t}(p_t)$ is resulting from another player's move.
-The states are the fruit of \textit{rival} actions, either from a human player or from another robot.
-So in terms of strategy optimization, the learner has to anticipate a possibly defavorable environment during the game and act accordingly.
-In particular, in this implementation, the learner plays after the rival, so it can win only with three or four moves.
-
-The signal function is set manually.
-It is triggered when a game ends.
-If the learner wins, it triggers $1$, $-0.1$ if there is a draw and $-1$ if it loses or if it does an illegal move.
-Notice that an illegal move corresponds only to playing on top of the other player.
-The learner here knows where it last played.
+The signal function is triggered when a game ends.
+If the learner wins, it returns $1$, $-0.1$ if there is a draw and $-1$ if the learner loses or if it does an illegal move.
+Notice that the latter corresponds only to playing on top of the other player as the learner knows where it last played.
 
@@
 Finally, I implemented an adaptive randomness factor. It has the very \textit{ad hoc} form 
-$$\epsilon = 30000/log_{10}(S)^6,$$
+$$\epsilon = 20000/log_{10}(S)^6,$$
 where $S$ is the size of the data in bytes.
-I chose this formula so it equals 99\% for a data corresponding to 10000 games which is how many games I train the learner with.
-The other robot is penalized with $\epsilon = 120000/log_{10}(S)^6$ so our learner has time to learn how to win in the beginning.
-This parameter has to be set sensibly for an optimized training to get a perfect "player".
-However we will not study this aspect further here.
+I chose this formula so it equals $\epsilon = 1\%$ after 10000 games, the number of games I trained the learner with.
+The other robot is penalized with $\epsilon = 300000/log_{10}(S)^6$ so our learner has time to experience wins in the beginning.
+This parameter has to be set sensibly for an optimized training to obtain a perfect player.
 
@@
 A few adjustments were to be made.
-First, notice that the learning phase is done against a Q-learning robot.
+First, as mentioned earlier, the learning phase is done against a Q-learning robot.
 After each of the players' move, an image representing the current state of the board is generated.
-This is image is then reduced and appended to the training set, as shown in Figure~\ref{fig:transition}.
+This is image is then reduced, as shown in Figure~\ref{fig:transition}, and appended to the training set.
 
@@
 \begin{center}
-\includegraphics[scale=.4]{"discounts"}
-\caption{Different discount rates affecting intermediary states}
+\includegraphics[scale=.4]{"transition"}
+\caption{Reduction of the original training images to grayscale 7x7 images}
 \label{fig:transition}
@@
 
-Second, our learner had the disadvantage of receiving imperfect information about the game compared to the other robot, knowing exactly where the moves are made.
-To compensate for this, our learner after each move rotates the game three times and appends the learned values accordingly.
+Second, our learner has the disadvantage of receiving imperfect information about the game compared to the other robot, knowing exactly where the moves are made.
+To compensate for this, our learner takes advantage of the equivalence between rotated states.
+After each move, it stores the four rotations the game in the training matrix.
 
-Third, as we do not have an objective sensor at our disposal, the signal recieved by the learner is the same as the signal coming from the oponent robot.
+Third, as we do not have an objective sensor at our disposal, the signal received by the learner is the same as the one coming from the opponent robot.
 It is an objective information indicating the final state of game when it happens.
 
-Third, the learning process, via neural networks, does not happen after each game but at regular intervals.
+Finally, the learning process, with the convolutional neural network, does not happen after each game but after a certain number of games.
 This is purely for computational constraints.
-However, this does not affect in fine the performance of the learner, which is able to win up to 90\% of the last phases' games.
+However, this does not affect \textit{in fine} the performance of the learner.
 
@@
 
-This section aims at assessing that the proposed algorithm leads to winning strategies.
+This section aims at assessing how this implementation of the proposed algorithm allows the learner to optimize its strategy.
 
 The following plots show the evolution the learner's performance during the learning phase.
-We also study how the proportions of wins and losses (and draws by deduction) vary according to the levels of $\epsilon$ for each type of robot.
-Recall that the robot updates its parameters every 100 games, and can potentially improve at this moment.
+We also study how the proportions of wins and losses, and draws by deduction, vary according to the levels of $\epsilon$ for each type of robot.
+Note that the robot updates its parameters every 500 games, and can potentially improve only after this.
 
-The green line corresponds to the proportion of won games for the last 100 games The red line corresponds to the proportion of lost games for the last 100 games.
+The blue line corresponds to the proportion of won games for the last 500 games. The red line corresponds to the proportion of lost games for the last 500 games.
 
-INCLUDE R PLOTS
+INCLUDE PLOTS
 
@@
 
-The proposed algorithm behaved as expected with its tic-tac-toe implementation.
-More work has to be done concerning the proof of convergence for the regressor function and the proof that asymptotically the behaviour of the learner is optimal.
+The tic-tac-toe player implementation provided a good demonstration for the proposed algorithm.
+However, more work has to be done concerning the proof of convergence for the approximation function and, more importantly, to prove the asymptotic gain-maximizing behavior of the learner.
 
-Now, we can think of more useful applications than this overkill tic-tac-toe player.
-For instance, a learner of this type, installed on augmented reality glasses could help people achieving precise tasks in a more optimal way.
-It could also be useful in online marketing to automatically identify set of actions leading to better sales, such as the sequence of clicks of customers.
-We also can imagine learning (thanks to vision) the way natural organisms evolve in their environment to build very adaptive machines.
+This work has been done considering recent advances in reinforcement learning and artificial neural networks.
+It also aims at anticipating a direction that applied machine learning could take in the next few years.
+I notably expect similar methods to have an impact in various industries and for private usages in the near future.
+
+For instance, such learning algorithm, adapted for augmented reality glasses could help people achieving precise tasks.
+We could even imagine such system to help doctors realize high precision surgery - recently a robot repaired a pig's bowel more accurately than a doctor.
+This would require a careful tuning of the reinforcement algorithm.
+And a way to improve their structure and its functions' parameters could be to observe the way natural organisms evolve in their environment.
+Such inspiration was a always a good guide to build agnostic, highly adaptive machines.
 
@@
 
-\bibitem{c1} Sutton and Barto, Reinforcement learning: An introduction, Vol. 2. No. 1, MIT press, 2012
+\bibitem{c1} Melo et al. (2008), An Analysis of Reinforcement Learning with Function Approximation,
 
-\bibitem{c2} $http://www0.cs.ucl.ac.uk/staff/D.Silver/web/Teaching_files/FA.pdf$
+\bibitem{c2} Mnih et al. (2013), Playing Atari with Deep Reinforcement Learning, arXiv:1312.5602 [cs.LG]
 
-\bibitem{c3} https://www.cs.toronto.edu/~vmnih/docs/dqn.pdf
+\bibitem{c3} Mnih et al. (2015) Human-level control through deep reinforcement learning, Nature 518, 529–533
 
-\bibitem{c4} http://gaips.inesc-id.pt/~fmelo/pub/melo08icml.pdf
+\bibitem{c4} Munos et al. (2016), Safe and efficient off-policy reinforcement learning, arXiv:1606.02647v1 [cs.LG]
 
-\bibitem{c5} Human-level control through deep reinforcement learning
+\bibitem{c5} Seijen et al. (2015), True Online Temporal-Difference Learning, arXiv:1512.04087v1 [cs.AI]
 
-\bibitem{c6} Munos et al. (2016), Safe and efficient off-policy reinforcement learning, arXiv:1606.02647v1 [cs.LG]
+\bibitem{c6} Shaul et al. (2015), Universal Value Function Approximators,  JMLR: W and CP volume 37
 
-\bibitem{c7} Seijen et al. (2015), True Online Temporal-Difference Learning, arXiv:1512.04087v1 [cs.AI]
+\bibitem{c7} Sutton and Barto, Reinforcement learning: An introduction, Vol. 2. No. 1, MIT press, 2012
 

### 2016-06-28 imgresize.py

@@ -1,7 +1,7 @@
 from PIL import Image, ImageEnhance
 import os
 os.chdir('/Users/Thomas/git/thesis/images7')
-dim = 3
+dim = 7
 
 def normalize(arr):
     arr = arr.astype('float')

### 2016-06-28 tic21.py

@@ -33,8 +33,8 @@ Ns = []
 maxi = 0
 pred = 0
 who = 'none'
-dim = 3
-selfplays = 1000
+dim = 7
+selfplays = 5000
 for i in os.listdir(smalls):
     if i.startswith('X'):
         Xs.append(scipy.misc.imread(i, flatten=True))
@@ -93,8 +93,10 @@ class Learner:
 
     def nn_train(s, X, y):
         n = X.shape[0]
-        cv_set = random.sample(range(0, n), int(round(n*0.2, 0)))
+        cv_set = random.sample(range(0, n), int(round(n*0.25, 0)))
         Xcv = X[cv_set,:] ; ycv = y[cv_set]
+        X = np.delete(X, cv_set, axis=0)
+        y = np.delete(y, cv_set)
         X = X.reshape(-1, 1, dim*3, dim*3)
         Xcv = Xcv.reshape(-1, 1, dim*3, dim*3)
         cv = (Xcv, ycv)
@@ -107,18 +109,17 @@ class Learner:
 
         s.net = Regressor(
                 layers=[
-                    Convolution("Rectifier", channels=27, kernel_shape=(3,3), kernel_stride=(2,2)),
-                    Convolution("Rectifier", channels=54, kernel_shape=(2,2), kernel_stride=(1,1)),
-                    #Convolution("Rectifier", channels=27, kernel_shape=(3,3), kernel_stride=(1,1)),
+                    Convolution("Rectifier", channels=441, kernel_shape=(7,7), kernel_stride=(7,7)),
+                    Layer("Rectifier", units=210),
                     Layer("Sigmoid", units=210),
                     Layer("Rectifier", units=210),
                     Layer("Sigmoid")],
                 n_iter=100,
-                learning_rate=0.001,
+                learning_rate=0.0001,
                 parameters=params,
                 valid_set=cv,
                 learning_rule='rmsprop',
-                f_stable=0.001,
+                f_stable=0.0001,
                 verbose=True,
                 batch_size=200,
                 n_stable=10)
@@ -225,7 +226,7 @@ class Learner:
             log_size = np.log10(data_size)
         else: log_size = 1
 
-        if s.player == 1: s.epsilon = min([1, 150000/(log_size**6)]) #we penalize the Q-learner. It would be too strong at the begining!
+        if s.player == 1: s.epsilon = min([1, 300000/(log_size**6)]) #we penalize the Q-learner. It would be too strong at the begining!
[…]
